### 2023-05-11 analysis/Manuscript/Paper_RFigures.Rmd
Initial commit of BenMAP input, output data, and analysis scripts

---
title: "Plot Figures for McDuffie et al., 2023 (those generated in R)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/CCD-CSIB/GitHub/Code/MOMM-RFT")
#setwd("~/CCD-CSIB/GitHub/Code/MOMM-RFT")
```

```{r libraries, include = FALSE}
library('tidyverse')
library('reshape2')
library('patchwork')
library('RColorBrewer')
library('maps')
library('mapproj')
library('arrow')
library('readxl')
library('zoo')
#library('Rfast')
```

# Figures

##Figure S1

Time series of global methane and O3 mixing ratios, by GCM
```{r FigureS1}

#specify input and output paths
Plots <-file.path(getwd(),'analysis','Manuscript','Plots')
Input <- file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs')

#Define BemMAP years and create dataframes
Model.years<-c(2020,2025,2030,2035,2040,2050,2060,2070,2080,2090,2100)
s1Data_methane <- data.frame(Model.years)
s1Data_ozone <- data.frame(Model.years)

#Read in the globally averaged mean O3 response data
# Generated by /analaysis/Manuscript/BenMAP/Inputs/Generate AQ Surfaces for BenMAP.R
All.mean.results <- read.csv(file.path(Input,"AQ Surfaces","Mean_Pulse_Data.csv"))

#pull out the methane data for each year (note, bad practice to reference by index - change later)
# the model name, and calculate the change in O3 between the baseline and perturbed scenarios
for (iyear in 1:length(Model.years)){
  s1Data_methane$MethanePerturbed_ppbv[iyear] <- All.mean.results[1,7+iyear]
  s1Data_ozone$CESM2[iyear]  <- All.mean.results[All.mean.results$Model =='CESM2',18+iyear]
  s1Data_ozone$GFDL[iyear]   <- All.mean.results[All.mean.results$Model =='GFDL',18+iyear]
  s1Data_ozone$GISS[iyear]   <- All.mean.results[All.mean.results$Model =='GISS',18+iyear]
  s1Data_ozone$HadGEM[iyear] <- All.mean.results[All.mean.results$Model =='HadGEM',18+iyear]
  s1Data_ozone$MIROC[iyear]  <- All.mean.results[All.mean.results$Model =='MIROC',18+iyear]
  s1Data_ozone$CESM2_delta[iyear] <- 
    s1Data_ozone$CESM2[iyear] - All.mean.results[All.mean.results$Model =='CESM2','Values.base'] #subtract baseline
  s1Data_ozone$GFDL_delta[iyear] <- 
    s1Data_ozone$GFDL[iyear] - All.mean.results[All.mean.results$Model =='GFDL','Values.base'] #subtract baseline
  s1Data_ozone$GISS_delta[iyear] <- 
    s1Data_ozone$GISS[iyear] - All.mean.results[All.mean.results$Model =='GISS','Values.base'] #subtract baseline
  s1Data_ozone$HadGEM_delta[iyear] <- 
    s1Data_ozone$HadGEM[iyear] - All.mean.results[All.mean.results$Model =='HadGEM','Values.base'] #subtract baseline
  s1Data_ozone$MIROC_delta[iyear] <- 
    s1Data_ozone$MIROC[iyear] - All.mean.results[All.mean.results$Model =='MIROC','Values.base'] #subtract baseline
}

s1Data_ozonelong <- reshape2::melt(s1Data_ozone, id="Model.years")

# Perturbed Methane timeseries
p1 <- ggplot(s1Data_methane,aes(x=Model.years,y=MethanePerturbed_ppbv, color = "Methane"))+
  geom_hline(aes(yintercept = 1834, linetype = "Baseline"), show.legend =FALSE)+
  geom_line()+
  geom_point()+
  scale_y_continuous(name="Methane Mixing Ratio [ppbv]")+
  labs(x='Year',y = "Methane Mixing Ratio [ppbv]", color="")+
  scale_x_continuous(breaks=seq(2020,2100,10))+
  ggtitle('Annual Methane')+
  theme_minimal()+
  theme(legend.position = 'bottom',
        axis.text.x=element_text(angle=45,hjust=1, vjust=1))+
  scale_color_manual(name='', values = c('black','black'))+
  scale_linetype_manual(name = '', values=c(Baseline = 'dashed'))
  #p1

#ozone perturbation with baseline
p2 <- s1Data_ozonelong %>%
  filter(!grepl('delta', variable)) %>%
  ggplot(aes(x=Model.years,y=value, color = variable))+
  scale_y_continuous(name="Ozone Mixing Ratio [ppbv]", limits = c(27, 31))+
  labs(x='Year',y = "Ozone Mixing Ratio [ppbv]", color="")+
  scale_x_continuous(breaks=seq(2020,2100,10))+
  geom_hline(aes(yintercept = All.mean.results$Values.base[1], linetype = "Baseline"), 
             show.legend =FALSE)+
  geom_hline(aes(yintercept = All.mean.results$Values.base[2], linetype = "Baseline"), 
             show.legend =FALSE)+
  geom_hline(aes(yintercept = All.mean.results$Values.base[3], linetype = "Baseline"), 
             show.legend =FALSE)+
  geom_hline(aes(yintercept = All.mean.results$Values.base[4], linetype = "Baseline"), 
             show.legend =FALSE)+
  geom_hline(aes(yintercept = All.mean.results$Values.base[5], linetype = "Baseline"), 
             show.legend =FALSE)+
  geom_line()+
  geom_point()+
  ggtitle('Summertime MDA8'~O[3])+
  theme_minimal()+
  theme(legend.position = 'bottom',
        axis.text.x=element_text(angle=45,hjust=1, vjust=1))+
  guides(color=guide_legend(ncol=3))+
  scale_colour_manual(values=c(brewer.pal(11,"Spectral")[c(1,4,9:11)]))+
  scale_linetype_manual(name = '', values=c(Baseline = 'dashed'))

#p2
p3<- p1+p2
p3 <-p3+plot_annotation(title = "Methane & Ozone Mixing Ratios in Response to a 2020 Pulse of Methane")

ggsave(plot = p3, filename = 
         file.path(Plots,'FigS1_wbackground.png'),width = 8,height = 5)
p3

#ozone perturbation without baseline
p2 <- s1Data_ozonelong %>%
  filter(grepl('delta', variable)) %>%
  mutate(value = value*1000) %>% #convert to pptv
  ggplot(aes(x=Model.years,y=value, color = variable))+
  scale_y_continuous(name="Ozone Mixing Ratio [pptv]", limits = c(0,400))+
  labs(x='Year',y = "Ozone Mixing Ratio [pptv]", color="")+
  scale_x_continuous(breaks=seq(2020,2100,10))+
  geom_hline(aes(yintercept = 0, linetype = "Baseline"), show.legend =FALSE)+
  geom_line()+
  geom_point()+
  ggtitle('Summertime Delta MDA8'~O[3])+
  theme_minimal()+
  theme(legend.position = 'bottom',
        axis.text.x=element_text(angle=45,hjust=1, vjust=1))+
  guides(color=guide_legend(ncol=3))+
  scale_colour_manual(values=c(brewer.pal(11,"Spectral")[c(1,4,9:11)]))+
  scale_linetype_manual(name = '', values=c(Baseline = 'dashed'))

p4<- p1+p2
p4 <-p4+plot_annotation(title = "Methane & Delta Ozone Mixing Ratios in Response to a 2020 Pulse of Methane")

ggsave(plot = p4, filename = 
         file.path(Plots,'FigS1.png'),width = 8,height = 5)
p4

```

##Figure S3

Time series of population characteristics 
```{r popcharacteristics}

#define paths
Plots <-file.path('analysis','Manuscript','Plots') 
inputs <- file.path("analysis","Manuscript")
inputs2 <- file.path('input','RFF','lookup_tables')
inputs3 <- file.path("input")
Inputs_socio <- file.path('input','RFF','rft_inputs')

#Read in RFF-SP data
# Pre-processed in /code/1_build_inputs_for_momm_rft.R
rffsp <- read_parquet(file.path(Inputs_socio, 'rffsp_pop_gdp_all_trials.parquet'))

rffsp_mean_pop <- rffsp %>%
  select(-c('dollar.year','gdp')) %>%
  group_by_at(.vars = c('Year','Country')) %>% #sum across trials for each country
  summarise_at(.vars = c('pop'), mean) %>%
  ungroup() %>%
  #interpolate between 5 year intervals
  group_by(Country) %>%
  complete(Year=min(Year):max(Year))%>%
  mutate(pop = approx(x=Year,y=pop,xout=2020:2100)$y)

rffsp_global_all_pop <- rffsp %>%
  group_by_at(.vars = c('Year','trial')) %>% #sum across countries
  summarise_at(.vars = c('pop'), sum) %>%
  ungroup() %>%
  group_by_at(.vars = c('trial')) %>%
  complete(Year=min(Year):max(Year))%>%
  mutate(pop = approx(x=Year,y=pop,xout=2020:2100)$y) %>%
  ungroup

p1 <-  rffsp_global_all_pop %>%
        group_by(Year) %>% 
        mutate(pop = pop/1e9) %>%
        summarize(X025=quantile(pop,probs=0.025), 
            X50=quantile(pop, probs=0.50),
            X975=quantile(pop,probs=0.975),
            X005=quantile(pop,probs=0.005),
            X995=quantile(pop,probs=0.995),
            mean=mean(pop)) %>% ungroup %>%
        na.omit() %>%
      ggplot()+
        geom_ribbon(aes(x=Year, 
                        ymin = X005, 
                        ymax = X995, 
                        fill="99th% Confidence Interval"), 
                    alpha=0.3) +
        geom_ribbon(aes(x=Year, 
                        ymin = X025, 
                        ymax = X975, 
                        fill = "95th% Confidence Interval"), 
                    alpha=0.9) +
        geom_line(aes(x=Year, y=mean,color="Mean"), size = 1)+
        scale_linetype_manual(values=c(rep("solid",5),"solid","longdash"))+
        scale_colour_manual("",values = c("95th% Confidence Interval" = 'darkgray',
                                        "99th% Confidence Interval" ='lightgray',
                                        "Mean" = "red"))+
        scale_fill_manual("",values = c("95th% Confidence Interval" = 'darkgray',
                                        "99th% Confidence Interval" ='lightgray',
                                        "Mean" = "tranparent"))+
        xlab("Year")+ylab("Population (billions)")+
        theme_minimal()+
        scale_x_continuous(breaks=seq(2020,2100,10))+
        theme(legend.position = c(0.3,0.9),
              legend.title=element_text(size=16), 
              legend.text=element_text(size=16),
              legend.spacing.y = unit(-0.45, 'cm'))+
        theme(axis.text = element_text(size=16),
              axis.title = element_text(size=16))#+
p1

ggsave(plot = p1, filename = 
         file.path(Plots,'Timeseries_global_population.png'),width = 8,height = 5)

####### Respiratory Mortality Rate & Population #############

#IF Data on ratios of respiratory to total mortality rates
IFsFactors<-read.csv(file.path(inputs,'IFP',"adjustment_factors.csv"))  
IFsCountry<-read.csv(file.path(inputs,'IFP',"country_codes.csv"))
  names(IFsCountry)[1] <- 'RegionName2'
IFsCtryCross<-read.csv(file.path(inputs,'IFP',"country_crosswalk.csv"))
IFsFactors<-left_join(IFsFactors,IFsCountry,by=c("Region"="RegionNum"))
IFsFactors<-left_join(IFsFactors,IFsCtryCross,by=c("RegionName2"="IFsName"),multiple='all')
  names(IFsFactors)<-c("X","Years","Region","RFF_Age_range","final_AF",
                       "RegionName","LocID","RFFName","Notes")
IFsFactors<-IFsFactors[,2:8]
Cross.year<-read_xlsx(file.path(inputs,'IFP','Crosswalk RFF Mort Year.xlsx'))
mean.pop<-read.csv(file.path(inputs,'RFF',"rff_pop_means.csv"))
#subset for relevant years (<2100)
mean.pop<-mean.pop[(mean.pop$Year<=2100),]
mean.pop <- mean.pop %>%
  filter(Year <= 2100) %>%
  mutate(mean_pop = mean_pop *1000) #convert to ppl counts

#calculate respiratory mortality counts (from rates * population)
mort_means <- read_csv(file.path(inputs,'RFF','rff_mort_means.csv'))
resp_mort_means <-left_join(mort_means,Cross.year,by=c("Year"="RFF year"))
resp_mort_means<-left_join(resp_mort_means,mean.pop,by=c("LocID","Min.Year"="Year","Age"))
resp_mort_means<-left_join(resp_mort_means,IFsFactors,by=c("LocID","Year"="Years",
                                                           "Age"="RFF_Age_range"))
#calculate mean resp mortality 
resp_mort_means$mean_resp_rate<-resp_mort_means$mean_mort_rate * resp_mort_means$final_AF
resp_mort_means<-resp_mort_means[(resp_mort_means$Min.Year<2100),]
resp_mort_means_noage<-resp_mort_means %>%
                       group_by(LocID,Min.Year) %>%
                       summarize(meanmort=weighted.mean(mean_resp_rate,mean_pop),.groups="keep") #population weighted rates
resp_mort_means_noage<-resp_mort_means_noage %>%
                        group_by(LocID) %>%
                        complete(Min.Year=min(Min.Year):max(Min.Year))%>%
                        mutate(MeanMortalityRate = na.approx(meanmort)) 
resp_mort_means_noage.2095<-resp_mort_means_noage[(resp_mort_means_noage$Min.Year==2095),]
resp_mort_means_noage.2095<-resp_mort_means_noage.2095 %>%
                                group_by(LocID) %>%
                                complete(Min.Year=c(2095:2100))%>%
                                fill(MeanMortalityRate,.direction="down")
#join 2095-2100 with rest of data (exclude 2095 from second table otherwise you'll have duplicates)
resp_mort_means_noage <-
  rbind(resp_mort_means_noage,
        resp_mort_means_noage.2095[(resp_mort_means_noage.2095$Min.Year>2095),])

#calculate global respiratory mortality counts for all scenarios
mort.files<-list.files(file.path(inputs2),
                       pattern = "All Trajectory Resp Mortality Ratios_AllAges_through 2100_")
All.mort<-lapply(file.path(file.path(inputs2),mort.files), readRDS)
resp_mort_ratios<-do.call(rbind,All.mort)
names(resp_mort_ratios)[5]<-"MortRatio"  #change Ratios to MortRatio so differs from pop column 
rm(mort.files,All.mort)
#it doesn't matter that the trajectory #s here are the mortality (not RFF) trajectory numbers. We are just taking the stats across all trials

resp_mort_ratios_comb <- left_join(resp_mort_ratios, 
                                   resp_mort_means_noage, 
                                   by = c("Year"="Min.Year","LocID"))

resp_mort_ratios_comb$MortRate = resp_mort_ratios_comb$MortRatio * 
  resp_mort_ratios_comb$MeanMortalityRate #get absolute rate for each trajectory

# #now need to calculate deaths to get global mortality rate
SampleID<-read.csv(file.path(inputs3,"sampled_pop_trajectory_numbers.csv"))
# #need to crosswalk RFF and mort trial #s
 Trajectory <- data.frame(rff_trial = c(1:10000),
                          mort_traj = SampleID$x)

Ctry.Name<-read.csv(file.path(inputs3,'Final Country Grid Col_Row Index.csv'))

 rffsp_mort <- rffsp %>%
   select('pop','Year','trial','Country') %>%
   left_join(Trajectory, by = c('trial'='rff_trial'))

 rffsp_mort2 <- rffsp_mort %>%
   select(-c(trial)) %>%
   distinct() %>%
   left_join(Ctry.Name, by = c('Country'='RFF_iso_code')) %>%
   select(c('pop','Year','mort_traj','COL')) %>%
   left_join(resp_mort_ratios_comb, by = c('COL'='LocID','Year','mort_traj'='Trajectory')) %>%
   select('Year','pop','mort_traj','COL','MortRate') #%>%

rffsp_mort3 <- rffsp_mort2 %>%
  mutate(mort = pop * MortRate) %>%
  group_by_at(.vars= c('Year','mort_traj')) %>%
  summarise_at(.vars= c("mort","pop"),sum, na.rm=T) %>% #sum across countries (not trials)
  ungroup %>%
  group_by(mort_traj) %>%
  complete(Year=min(Year):max(Year))%>%
  mutate(pop = approx(x=Year,y=pop,xout=2020:2100)$y,
         mort = approx(x=Year,y=mort,xout=2020:2100)$y) %>%
  mutate(global_rate = mort/pop)

p2 <-  rffsp_mort3 %>%
        group_by(Year) %>% 
        mutate(MortRate = global_rate*1e5) %>% #get per 100K rate
        summarize(X025=quantile(MortRate,probs=0.025), 
            X50=quantile(MortRate, probs=0.50),
            X975=quantile(MortRate,probs=0.975),
            X005=quantile(MortRate,probs=0.005),
            X995=quantile(MortRate,probs=0.995),
            mean=mean(MortRate)) %>% ungroup %>%
        na.omit() %>%
      ggplot()+
        geom_ribbon(aes(x=Year,
                        ymin = X005, 
                        ymax = X995, 
                        fill="99th% Confidence Interval"), 
                    alpha=0.3) +
        geom_ribbon(aes(x=Year, 
                        ymin = X025, 
                        ymax = X975, 
                        fill = "95th% Confidence Interval"), 
                    alpha=0.9) +
        geom_line(aes(x=Year, y=mean,color="Mean"), size = 1)+
        scale_linetype_manual(values=c(rep("solid",5),"solid","longdash"))+
        scale_colour_manual("",values = c("95th% Confidence Interval" = 'darkgray',
                                        "99th% Confidence Interval" ='lightgray',
                                        "Mean" = "red"))+
        scale_fill_manual("",values = c("95th% Confidence Interval" = 'darkgray',
                                        "99th% Confidence Interval" ='lightgray',
                                        "Mean" = "tranparent"))+
        xlab("Year")+ylab("Respiratory Mortality Rate (per 100K)")+
        theme_minimal()+
        scale_x_continuous(breaks=seq(2020,2100,10))+
        theme(legend.position = c(0.3,0.9),
              legend.title=element_text(size=16), 
              legend.text=element_text(size=16),
              legend.spacing.y = unit(-0.45, 'cm'))+
        theme(axis.text = element_text(size=16),
              axis.title = element_text(size=16))#+
p2

ggsave(plot = p2, filename = 
         file.path(Plots,'Timeseries_global_avgMortRate.png'),width = 8,height = 5)

p3<- p1+p2
p3 <-p3+plot_annotation(title = "Global Population & Mortality Rate Projections",
                        theme = theme(plot.title = element_text(size = 20)))

ggsave(plot = p3, filename = 
         file.path(Plots,'FigS3.png'),width = 14,height = 5)
p3

```

##Figures 2a, S4

Timeseries of Respiratory mortality counts, by GCM (Figure S4)
Time series of respiratory mortality, with beta uncertainties (Fig 2a)
```{r Global & Country TimeSeries}

#plot line graph of global total and top 10 countries (for MMM)
Plots <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')

# Processed BenMAP respiratory mortality output files
# from /analysis/Manuscript/BenMAP/Summarizing BenMAP Results.R
countryResults    <- read.csv(paste0(Plots,"/Country Results Summary by Model & Year.csv"))
globalResults     <- read.csv(paste0(Plots,"/Global Results Summary by Model & Year.csv"))
globalResults2.5  <- read.csv(paste0(Plots,"/Global Results Summary by Model & Year 2_5CI.csv"))
globalResults97.5 <- read.csv(paste0(Plots,"/Global Results Summary by Model & Year 97_5CI.csv"))

globalResultslong <- globalResults %>%
  select(-c('AvgResponse','TotalResults','TotalPct2_5','TotalPct97_5'))
globalResultslong <- reshape2::melt(globalResultslong, id="Model")
globalResultslong$Model.years <- strtoi(str_replace(globalResultslong$variable, 'X',''))

###### do same calcs for CIs
globalResults2.5long <- globalResults2.5 %>%
  select(-c('AvgResponse','TotalResults','TotalPct2_5','TotalPct97_5'))
globalResults2.5long <- reshape2::melt(globalResults2.5long, id="Model")
globalResults2.5long$Model.years <- strtoi(str_replace(globalResults2.5long$variable, 'X',''))

globalResults97.5long <- globalResults97.5 %>%
  select(-c('AvgResponse','TotalResults','TotalPct2_5','TotalPct97_5'))
globalResults97.5long <- reshape2::melt(globalResults97.5long, id="Model")
globalResults97.5long$Model.years <- strtoi(str_replace(globalResults97.5long$variable, 'X',''))

######
### Figure S4.

p1 <-  ggplot(globalResultslong, aes(x=Model.years,y=value,colour=Model,group=Model,linetype=Model))+
        geom_point()+
        geom_line()+
        scale_linetype_manual(values=c(rep("solid",5),"solid","longdash"))+
        scale_colour_manual(values=c(brewer.pal(11,"Spectral")[c(1,4,9:11)],"black","gray"))+
        scale_size_manual(values=c(rep(1,5),1.5,1))+
        xlab("Year")+ylab("Annual Attributable Respiratory-Related Mortality (deaths)")+
        ggtitle('Annual Respiratory-Related Mortality Attributable to an Increase in \nMDA8 Ozone Exposure from a Methane Pulse in 2020')+
        theme_minimal()+
        scale_x_continuous(breaks=seq(2020,2100,10))+
        scale_y_continuous(limits=c(0,18000))
p1
ggsave(plot = p1, filename = 
         file.path(Plots,'Plots/Timeseries_O3Mort_GCMs.png'),width = 8,height = 5)

####

# ### Extra Fig for now ###
# #plot global MMM with uncertainty bounds (from GCM uncertainty)
# global_limits<- globalResultslong %>%
#   filter(Model %in% c('MMM - Mean AQS','MIROC','HadGEM')) %>%
#   spread(Model,value)
# p1_bounds <- global_limits %>%
#     ggplot(aes(x=Model.years,y=`MMM - Mean AQS`,ymin=MIROC, ymax=HadGEM,fill='MMM &\nGCM Distribution'))+
#         geom_point()+
#         geom_line()+
#         geom_ribbon(alpha=0.5) +
#         scale_linetype_manual(values="solid","longdash","longdash")+
#         scale_colour_manual('',values="black")+
#         scale_fill_manual("",values="gray50")+
#         scale_size_manual(values=c(rep(1,5),1.5,1))+
#         xlab("Year")+ylab("Deaths")+
#         ggtitle('Global Annual Respiratory-Related Ozone Attributable Mortality')+
#         theme_minimal()+
#         scale_x_continuous(breaks=seq(2020,2100,10))+
#         scale_y_continuous(limits=c(0,18000), labels=scales::comma)+
#         theme(legend.position = c(0.15, 0.15), legend.text = element_text(size=16))+
#         theme(axis.text = element_text(size=16),
#               axis.title = element_text(size=16),
#               axis.text.x=element_text(angle=45,hjust=1, vjust=1))
# p1_bounds
# ggsave(plot = p1_bounds, filename = 
#          file.path(Plots,'Plots/Timeseries_O3Mort_GCMbounds.png'),width = 9,height = 5)

## Figure 2a

#plot global MMM with uncertainty bounds (from CRF uncertainty)
global_limits<- globalResults2.5long %>%
  filter(Model %in% c('MMM - Mean AQS')) %>%
  mutate(Model = 'lower') %>%
  rbind(globalResults97.5long) %>%
  filter(Model %in% c('lower','MMM - Mean AQS')) %>%
  mutate(Model = case_when( Model == 'MMM - Mean AQS' ~ 'upper',TRUE ~ Model)) %>%
  #spread(Model,value) %>%
  #rename(upper = `MMM - Mean AQS`) %>%
  rbind(globalResultslong) %>%
  filter(Model %in% c('lower','upper','MMM - Mean AQS')) %>%
  spread(Model,value)
p1_bounds <- global_limits %>%
    ggplot(aes(x=Model.years,y=`MMM - Mean AQS`,ymin=lower, ymax=upper,fill='MMM &\nCRF Distribution'))+
        geom_point()+
        geom_line()+
        geom_ribbon(alpha=0.5) +
        scale_linetype_manual(values="solid","longdash","longdash")+
        scale_colour_manual('',values="black")+
        scale_fill_manual("",values="gray50")+
        scale_size_manual(values=c(rep(1,5),1.5,1))+
        xlab("Year")+ylab("Deaths")+
        ggtitle('Global Annual Respiratory-Related Ozone Attributable Mortality')+
        theme_minimal()+
        scale_x_continuous(breaks=seq(2020,2100,10))+
        scale_y_continuous(limits=c(0,18000), labels=scales::comma)+
        theme(legend.position = c(0.15, 0.15), legend.text = element_text(size=16))+
        theme(axis.text = element_text(size=16),
              axis.title = element_text(size=16),
              axis.text.x=element_text(angle=45,hjust=1, vjust=1))
p1_bounds
ggsave(plot = p1_bounds, filename = 
         file.path(Plots,'Plots/Timeseries_O3Mort_CRFbounds.png'),width = 9,height = 5)

###

# ### Extra
# ## Line plot timeseries of deaths in top 10 countries.
# cResultslong <- countryResults %>%
#   arrange(desc(TotalResults)) %>%
#   group_by(Model) %>%
#   slice(1:10) %>%
#   ungroup()%>%
#   filter(Model =='MMM') %>%
#   select(-c('AvgResponse','TotalResults','TotalPct2_5','TotalPct97_5','Model','Region','SuperRegion')) 
# 
# cResultslong <- reshape2::melt(cResultslong, id="COUNTRY")
# cResultslong$Model.years <- strtoi(str_replace(cResultslong$variable, 'X',''))
# globalMMM <- globalResultslong[globalResultslong$Model == 'MMM - Mean AQS',]
# globalMMM$COUNTRY <- 'Global'
# globalMMM <- globalMMM %>%
#   select(-c('Model'))
# cResultslong <- rbind(cResultslong, globalMMM)
#     
#   cou_i <- cResultslong %>% 
#   group_by_at(.vars=c("COUNTRY")) %>% 
#   #filter(Model.years == 2020) %>%
#   summarize_at(.vars=c("value"),sum,na.rm=T) %>%
#   ungroup %>% arrange(desc(value)) %>%
#   distinct(COUNTRY)
# 
#   cResultslong <- cResultslong %>% 
#   mutate(COUNTRY = COUNTRY %>% factor(levels = (cou_i$COUNTRY))) %>%
#   arrange_at(.vars=c("COUNTRY"))
# 
# p2 <-  ggplot(cResultslong, aes(x=Model.years,y=value,colour=COUNTRY,group=COUNTRY,linetype=COUNTRY))+
#         geom_point()+
#         geom_line()+
#         scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
#         scale_colour_manual(values=c("black",brewer.pal(11,"Spectral")[c(1:3,5,7:11)],"gray"))+
#         scale_size_manual(values=c(rep(1,5),1.5,1))+
#         xlab("Year")+ylab("Annual Attributable Respiratory-Related Mortality (deaths)")+
#         ggtitle('Annual Respiratory-Related Ozone Attributable Mortality (MMM), \nTop 10 Countries')+
#         theme_minimal()+
#         scale_x_continuous(breaks=seq(2020,2100,10))+
#         scale_y_continuous(limits=c(0,11500))
# p2
# 
# ## Extra
# ### Line plot timeseries of mortality by super-region
# 
# #Plot Regional Results
# #calculate average regional response
# #regional_avg <- countryResults %>%
# #  group_by_at(.vars = c('Model','SuperRegion')) %>%
# #  summarise_at(3,mean,na.rm=T) %>%
# #  ungroup() #%>%
# 
# cResultslong <- countryResults %>%
#   group_by_at(.vars = c('Model','SuperRegion')) %>%
#   summarise_at(4:17,sum,na.rm=T) %>%
#   ungroup() %>%
#   arrange(desc(TotalResults)) %>%
#   group_by(Model) %>%
#  # slice(1:10) %>%
#   ungroup()%>%
#   filter(Model =='MMM') %>%
#   select(-c('TotalResults','TotalPct2_5','TotalPct97_5','Model')) 
# 
# cResultslong <- reshape2::melt(cResultslong, id="SuperRegion")
# cResultslong$Model.years <- strtoi(str_replace(cResultslong$variable, 'X',''))
# globalMMM <- globalResultslong[globalResultslong$Model == 'MMM - Mean AQS',]
# globalMMM$SuperRegion <- 'Global'
# globalMMM <- globalMMM %>%
#   select(-c('Model'))
# cResultslong <- rbind(cResultslong, globalMMM)
# 
# #cResultslong <- cResultslong %>%
# #  group_by_at(c("COUNTRY","Model.years")) %>%
# #  arrange(desc(value)) %>%
# #  mutate(COUNTRY = factor(COUNTRY, levels=c(
#     
#   reg_i <- cResultslong %>% 
#   group_by_at(.vars=c("SuperRegion")) %>% 
#   #filter(Model.years == 2020) %>%
#   summarize_at(.vars=c("value"),sum,na.rm=T) %>%
#   ungroup %>% arrange(desc(value)) %>%
#   distinct(SuperRegion)
# 
#   cResultslong <- cResultslong %>% 
#   mutate(SuperRegion = SuperRegion %>% factor(levels = (reg_i$SuperRegion))) %>%
#   arrange_at(.vars=c("SuperRegion"))
# 
# p3 <-  ggplot(cResultslong, aes(x=Model.years,
#                                 y=value,
#                                 colour=SuperRegion,
#                                 group=SuperRegion,
#                                 linetype=SuperRegion))+
#         geom_point()+
#         geom_line()+
#         scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
#         scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
#         scale_size_manual(values=c(rep(1,5),1.5,1))+
#         xlab("Year")+ylab("Annual Attributable Respiratory-Related Mortality (deaths)")+
#         ggtitle('Annual Respiratory-Related Ozone Attributable Mortality (MMM), \nby GBD Super Region')+
#         theme_minimal()+
#         scale_x_continuous(breaks=seq(2020,2100,10))+
#         scale_y_continuous(limits=c(0,11500))
# p3
# 
# 
# 
# ### Extra 
# # stacked area chart or mortality by super region
# regsum <- countryResults %>%
#   filter(Model == 'MMM')%>%
#   select(SuperRegion, TotalResults) %>%
#   group_by_at(.vars = c('SuperRegion')) %>%
#   summarise_at(.vars=c("TotalResults"),sum,na.rm=F) %>% ungroup
# 
# regresp <- countryResults %>%
#   filter(Model == 'MMM')%>%
#   select(SuperRegion, AvgResponse) %>%
#   group_by_at(.vars = c('SuperRegion')) %>%
#   summarise_at(.vars=c("AvgResponse"),mean,na.rm=F) %>% ungroup
# 
# 
# 
# p4 <-  cResultslong %>%
#         filter(SuperRegion!='Global')%>%
#         ggplot( aes(x=Model.years,
#                                 y=value,
#                                 fill=SuperRegion,
#                                 group=SuperRegion))+
#         #geom_point()+
#         geom_area()+
#         #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
#         scale_fill_manual('GBD Super Regions',
#                           values=c(brewer.pal(7,"Spectral")))+
#         scale_size_manual(values=c(rep(1,5),1.5,1))+
#         xlab("Year")+ylab("Deaths")+
#         ggtitle('Annual Respiratory-Related Ozone Attributable Mortality (MMM), \nby GBD Super Region')+
#         theme_minimal()+
#         theme(legend.position = 'right')+ #c(0.7, 0.7))+
#         theme(axis.text = element_text(size=12),
#               axis.title = element_text(size=12),
#               axis.text.x=element_text(angle=45,hjust=1, vjust=1))+
#         scale_x_continuous(breaks=seq(2020,2100,10))+
#         scale_y_continuous(limits=c(0,11500), label = scales::comma)
#   
# p4
# ggsave(plot = p4, filename = 
#          file.path(Plots,'Timeseries_O3Mort_SuperRegion.png'),width = 10,height = 5)

```

```{r, countrymap}

#should these be per capita?
# add per capita code

## FIgure 2b and adapted to S5

#map of country mortality results
MODEL = 'HadGEM'

Plots <-file.path(getwd(),'Results','Global BenMAP WebTool')

countryResults <- read.csv(paste0(Plots,"/Country Results Summary by Model & Year_eem.csv"))

country_total <- countryResults %>%
  select(c('COUNTRY','Region','SuperRegion','Model','TotalResults','TotalPct2_5','TotalPct97_5')) %>%
  filter(Model == MODEL) %>%
  rename(region = COUNTRY) #%>%

#plot by super region?

#need to sum all country results

#map data
worldData <- map_data("world")
#correct region names
worldData["region"][worldData["region"] == "Russia"] <- "Russian Federation"
worldData["region"][worldData["region"] == "USA"] <- "United States"
worldData["region"][worldData["region"] == "Virgin Islands" & worldData['subregion']==' US'] <- "US Virgin Islands"
worldData["region"][worldData["region"] == "Taiwan"] <- "China, Taiwan Province of China"
worldData["region"][worldData["region"] == "Trinidad" | worldData['region']=='Tobago'] <- "Trinidad and Tobago"
worldData["region"][worldData["region"] == "Grenadines"| worldData['region']=='Saint Vincent'] <- "Saint Vincent and the Grenadines"
worldData["region"][worldData["region"] == "Swaziland"] <- "Eswatini"
worldData["region"][worldData["region"] == "UK"] <- "United Kingdom"
worldData["region"][worldData["region"] == "Cape Verde"] <- "Cabo Verde"
worldData["region"][worldData["region"] == "Republic of Congo"] <- "Congo"
worldData["region"][worldData["region"] == "Democratic Republic of the Congo"] <- "Congo DRC"
worldData["region"][worldData["region"] == "Ivory Coast"] <- "Cote d'Ivoire"
worldData["region"][worldData["region"] == "Brunei"] <- "Brunei Darussalam"
worldData["region"][worldData["region"] == "Antigua"| worldData['region']=='Barbuda'] <- "Antigua and Barbuda"
worldData["region"][worldData["region"] == "Palestine"] <- "Palestinian Territory"
worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Hong Kong'] <- "China, Hong Kong SAR"
worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Macao'] <- "China, Macao SAR"
worldData["region"][worldData["region"] == "Guernsey"| worldData['region']=='Jersey'] <- "Channel Islands"

#Merge data
MergedData <- left_join(worldData, country_total, by = 'region')
#replace zero with na
MergedData$TotalResults[MergedData$TotalResults ==0] <- NA

#chloropleth of world
p5 <- ggplot() +
  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = TotalResults),
               color = 'black',linewidth=0.15)+

#  scale_fill_binned("Total Deaths",
#                    breaks= c(0,100,1000,10000,80000),
#                    guide= guide_colorsteps(even.steps = FALSE))+
  scale_fill_fermenter("Total Deaths",
                       n.breaks=9,
                        type = 'seq',
                       direction = -1,
                        palette = "Spectral",
                        na.value='lightgray',
                       limits=c(0,100000),
                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
                       breaks= c(0,100,1000,10000,100000),
                       #n.breaks = 6,
                       label=scales::comma)+#,
                       #guide =  guide_colorbar(nbin=10))+

                       #guides() (fill= guide_colorbar(even.steps) = FALSE
  #scale_fill_continuous(name="Total Mortality (deaths per 100K)",
  #                      low="darkblue",
  #                      high="lightblue",
  #                      limits=c(0,70),
  #                      na.value='gray')+
  theme_void()+
  theme(legend.spacing.y = unit(.5,'cm'),
        legend.text = element_text(size=12),
        legend.title = element_text(size=13))+
  #guide(c)

  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))

# Map of the world:
#p5<- map('world',col="grey", fill=TRUE, bg="white", lwd=0.05, mar=rep(0,4),border=0, ylim=c(-80,80) )

p5
ggsave(plot = p5, filename = 
         file.path(Plots,'Plots',paste0('Total_MortMap_',MODEL,'.png')),width = 9,height = 5)
```

```{r, countrymap_percapita}

#should these be per capita?
# add per capita code

## Figure 2b and adapted to S5

#map of country mortality results
MODEL = 'MMM'

Plots <-file.path(getwd(),'Results','Global BenMAP WebTool')

countryResults <- read.csv(paste0(Plots,"/Country Results Summary by Model & Year_eem.csv"))
country_2020 <- countryResults %>%
  select(c('COUNTRY','Region','SuperRegion','Model','AvgResponse','X2020')) %>%
  filter(Model == MODEL) %>%
  rename(region = COUNTRY) #%>%

#read in population
inputs <- file.path("Code","InputData","Creating RFF Lookup Tables")
mean.pop<-read.csv(file.path(inputs,"rff_pop_means.csv")) #this dataset includes all countries (rff-sp does not)
mean.pop <- mean.pop %>%
  filter(Year ==2020) %>%
  mutate(mean_pop = mean_pop *1000) %>% #convert to ppl counts
  group_by(LocID) %>%
  summarise_at(.vars = c('mean_pop'), sum)

inputs3 <- file.path("Code","InputData","Tool Inputs")
Ctry.Name<-read.csv(file.path(inputs3,'Final Country Grid Col_Row Index_EEM.csv'))

mean.pop <- mean.pop %>% left_join(Ctry.Name, by = c('LocID'='COL')) %>%
  select(-c('Region','SuperRegion'))
country_plot1 <- country_2020 %>% left_join(mean.pop, by =c('region'='COUNTRY')) %>%
  mutate(death_percap = 1e5*X2020/mean_pop) #absolute deaths per 100K

#read in mort data (created in the mortality timeseries chunk)
mort_rate <- resp_mort_means_noage %>% 
  filter(Min.Year ==2020)
country_plot <- country_plot1 %>% left_join(mort_rate, by ='LocID')

#plot by super region?

#need to sum all country results

#map data
worldData <- map_data("world")
#correct region names
worldData["region"][worldData["region"] == "Russia"] <- "Russian Federation"
worldData["region"][worldData["region"] == "USA"] <- "United States"
worldData["region"][worldData["region"] == "Virgin Islands" & worldData['subregion']==' US'] <- "US Virgin Islands"
worldData["region"][worldData["region"] == "Taiwan"] <- "China, Taiwan Province of China"
worldData["region"][worldData["region"] == "Trinidad" | worldData['region']=='Tobago'] <- "Trinidad and Tobago"
worldData["region"][worldData["region"] == "Grenadines"| worldData['region']=='Saint Vincent'] <- "Saint Vincent and the Grenadines"
worldData["region"][worldData["region"] == "Swaziland"] <- "Eswatini"
worldData["region"][worldData["region"] == "UK"] <- "United Kingdom"
worldData["region"][worldData["region"] == "Cape Verde"] <- "Cabo Verde"
worldData["region"][worldData["region"] == "Republic of Congo"] <- "Congo"
worldData["region"][worldData["region"] == "Democratic Republic of the Congo"] <- "Congo DRC"
worldData["region"][worldData["region"] == "Ivory Coast"] <- "Cote d'Ivoire"
worldData["region"][worldData["region"] == "Brunei"] <- "Brunei Darussalam"
worldData["region"][worldData["region"] == "Antigua"| worldData['region']=='Barbuda'] <- "Antigua and Barbuda"
worldData["region"][worldData["region"] == "Palestine"] <- "Palestinian Territory"
worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Hong Kong'] <- "China, Hong Kong SAR"
worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Macao'] <- "China, Macao SAR"
worldData["region"][worldData["region"] == "Guernsey"| worldData['region']=='Jersey'] <- "Channel Islands"

#Merge data
MergedData <- left_join(worldData, country_plot, by = 'region')
#replace zero with na
MergedData$death_percap[MergedData$death_percap ==0] <- NA
MergedData$AvgResponse[MergedData$AvgResponse ==0] <- NA
MergedData$MeanMortalityRate[MergedData$MeanMortalityRate ==0] <- NA
MergedData <- MergedData %>%
  mutate(rel_death_percap = death_percap/mean(death_percap, na.rm=TRUE),
         rel_AvgResponse = AvgResponse/mean(AvgResponse, na.rm=TRUE),
         rel_mean_pop = mean_pop/mean(mean_pop, na.rm=TRUE),
         rel_MeanMortalityRate = MeanMortalityRate/mean(MeanMortalityRate, na.rm=TRUE))

#chloropleth of world
# Percapita deaths
p5 <- ggplot() +
  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = death_percap),
               color = 'black',linewidth=0.15)+
#  scale_fill_gradientn(
#   colors=c("red","white","blue"),
#   values=rescale(c(-1,0,1)),
#   limits=c(-1,1))+
  scale_fill_fermenter("Deaths Per Capita",
                       n.breaks=10.5,
                        type = 'seq',
                       direction = -1,
                        palette = "Spectral",
                        na.value='lightgray',
                       #values=rescale(c(-1,0,1)),
                       limits=c(#0,
                                min(MergedData$death_percap, na.rm=TRUE),
                                0.3),
                                #max(MergedData$death_percap, na.rm=TRUE)),
                       #midpoint=1,
                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
                       #breaks= c(0,100,1000,10000,100000),
                       #n.breaks = 6,
                       #label=scales::comma)+#,
                       labels = c('Minimum',rep('',8),'Maximum'))+
                       #guide =  guide_colorbar(nbin=10))+

                       #guides() (fill= guide_colorbar(even.steps) = FALSE
  #scale_fill_continuous(name="Total Mortality (deaths per 100K)",
  #                      low="darkblue",
  #                      high="lightblue",
  #                      limits=c(0,70),
  #                      na.value='gray')+
  theme_void()+
  theme(legend.position = 'right')+
  theme(legend.key.height = unit(0.5,'cm'),
        legend.spacing.y = unit(.5,'cm'),
        legend.text = element_text(size=12),
        legend.title = element_text(size=13))+
  #ggtitle('Total Deaths Per 100K')+
  #theme(plot.title = element_text(size = 20))+
  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))

# Map of the world:
#p5<- map('world',col="grey", fill=TRUE, bg="white", lwd=0.05, mar=rep(0,4),border=0, ylim=c(-80,80) )

p5

#mortality rate
p6 <- ggplot() +
  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, 
                                      fill = MeanMortalityRate),
               color = 'black',linewidth=0.15)+
  scale_fill_fermenter("Respiratory Mortality Rates",
                       n.breaks=9,
                        type = 'seq',
                       direction = -1,
                        palette = "Spectral",
                        na.value='lightgray',
                       limits=c(min(MergedData$MeanMortalityRate, na.rm=TRUE),
                                max(MergedData$MeanMortalityRate, na.rm=TRUE)),
                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
                       #breaks= c(0,100,1000,10000,100000),
                       #n.breaks = 6,
                       #label=scales::comma)+
                       labels = c('Minimum','','','','','','Maximum'))+

  theme_void()+
  ggtitle("b) Respiratory Mortality Rates Per 100K")+
  theme(legend.position = 'none')+
  theme(legend.spacing.y = unit(.5,'cm'),
        legend.text = element_text(size=12),
        legend.title = element_text(size=13))+
  theme(plot.title = element_text(size = 20))+
  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
p6

#population
p7 <- ggplot() +
  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = mean_pop),
               color = 'black',linewidth=0.15)+
  scale_fill_fermenter("Population",
                       n.breaks=9,
                        type = 'seq',
                       direction = -1,
                        palette = "Spectral",
                        na.value='lightgray',
                       limits=c(min(MergedData$mean_pop, na.rm=TRUE),
                                max(MergedData$mean_pop, na.rm=TRUE)),
                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
                       #breaks= c(0,100,1000,10000,100000),
                       #n.breaks = 6,
                       #label=scales::comma)+
                       labels = c('Minimum','','','','','','Maximum'))+
  theme_void()+
  ggtitle('a) Population')+
  theme(legend.position = 'none')+
  theme(legend.spacing.y = unit(.5,'cm'),
        legend.text = element_text(size=12),
        legend.title = element_text(size=13))+
  theme(plot.title = element_text(size = 20))+
  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
p7

#avg response
p8 <- ggplot() +
  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = AvgResponse),
               color = 'black',linewidth=0.15)+
  scale_fill_fermenter("",
                       n.breaks=9,
                        type = 'seq',
                       direction = -1,
                        palette = "Spectral",
                        na.value='lightgray',
                       limits=c(min(MergedData$AvgResponse, na.rm=TRUE),
                                max(MergedData$AvgResponse, na.rm=TRUE)),
                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
                       #breaks= c(0,100,1000,10000,100000),
                       #n.breaks = 6,
                       #label=scales::comma)+
                       labels = c('Minimum','','','','','','Maximum'))+
  theme_void()+
  ggtitle(expression(paste('c) Average Ozone Response (pptv ', O[3],"/", "ppbv ",CH[4],')')))+
  theme(legend.position = 'bottom')+
  theme(legend.key.width = unit(2,'cm'),
        legend.text = element_text(size=12),
        legend.title = element_text(size=13))+
  theme(plot.title = element_text(size = 20))+
  #guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
p8

p<- p7+p6+p8
p <-p+plot_annotation(title = "",
                        theme = theme(plot.title = element_text(size = 20)))+
      plot_layout(ncol = 1)
p

ggsave(plot = p, filename = 
         file.path(Plots,'Plots',paste0('Relative_2020_CombMaps_',MODEL,'.png')),width = 9,height = 15)

ggsave(plot = p5, filename = 
         file.path(Plots,'Plots',paste0('Relative_2020_DeathsPerCapMaps_',MODEL,'.png')),width = 9,height = 5)
p5
```

```{r, damagesbar}

#create bar graph of damages by region

## FIgure 2c

countryresults <-
  #read_csv(file.path('Results',"Country Results Interpolated Mortality Annual Undiscounted Damages.csv"),
   #        col_types = cols())
   read_csv(file.path('Results',"npd_country_damage_transformation_NoCessation.csv"),
           col_types = cols())

globalresults <-
  read_csv(file.path('Results',"npd_global_damage_transformation_NoCessation.csv"),
           col_types = cols())

globalresults <- 
  globalresults  %>%
  group_by_at(.vars = c('Model','discount.rate')) %>%
  arrange(desc(mean)) %>%
  filter(Model =='MMM') %>%
  mutate(SuperRegion = 'Global')%>%
  filter(discount.rate %in% 
           c('1.5% Ramsey', '2.0% Ramsey', '2.5% Ramsey', "3.0% Ramsey",'2.0% CDR','3.0% CDR'))

cResultslong <- countryresults %>%
  group_by_at(.vars = c('Model','SuperRegion','discount.rate')) %>%
  summarise_at(c('mean','2.5%','97.5%'),sum,na.rm=T) %>%
  ungroup() %>%
  arrange(desc(mean)) %>%
  #mutate(damages = damages/275e6,
  #       damages_2_5 = damages_2_5/275e6,
  #       damages_97_5 = damages_97_5/275e6) %>% #convert to per mt CH4 (not discounted)
  filter(Model =='MMM') %>%
  #select(-c('TotalResults','TotalPct2_5','TotalPct97_5','Model')) %>%
    rbind(globalresults)

## ordering of central rate for alpha transparency
cResultslong$discount.rate <- factor(cResultslong$discount.rate, 
                                  levels=c('3.0% CDR','2.0% CDR',
                                          "3.0% Ramsey","2.5% Ramsey", 
                                          "2.0% Ramsey", "1.5% Ramsey" ))

## function for labels
means = cResultslong %>% 
  group_by_at(c('discount.rate','SuperRegion')) %>% 
  summarise(means=mean) %>% 
  ungroup() %>%
  mutate(across(c('means'), ~ signif(.,3)))

lower = cResultslong %>% 
  group_by_at(c('discount.rate','SuperRegion')) %>% 
  summarise(means=`2.5%`) %>% 
  ungroup() %>%
  mutate(across(c('means'), ~ signif(.,3)))

upper = cResultslong %>% 
  group_by_at(c('discount.rate','SuperRegion')) %>% 
  summarise(means=`97.5%`) %>% 
  ungroup() %>%
  mutate(across(c('means'), ~ signif(.,3)))

# cResultslong <- reshape2::melt(cResultslong, id="SuperRegion")
# cResultslong$Model.years <- strtoi(str_replace(cResultslong$variable, 'X',''))
# globalMMM <- globalResultslong[globalResultslong$Model == 'MMM - Mean AQS',]
# globalMMM$SuperRegion <- 'Global'
# globalMMM <- globalMMM %>%
#   select(-c('Model'))
# cResultslong <- rbind(cResultslong, globalMMM)

#cResultslong <- cResultslong %>%
#  group_by_at(c("COUNTRY","Model.years")) %>%
#  arrange(desc(value)) %>%
#  mutate(COUNTRY = factor(COUNTRY, levels=c(

  reg_i <- cResultslong %>% 
  group_by_at(.vars=c("SuperRegion")) %>% 
  #filter(Model.years == 2020) %>%
  summarize_at(.vars=c("mean"),sum,na.rm=T) %>%
  ungroup %>% arrange((mean)) %>%
  distinct(SuperRegion)

  cResultslong <- cResultslong %>% 
  mutate(SuperRegion = SuperRegion %>% factor(levels = (reg_i$SuperRegion))) %>%
  arrange_at(.vars=c("SuperRegion"))

  p1 <-  ggplot(cResultslong, aes(x=SuperRegion, y=mean))+
        geom_linerange(aes(x=SuperRegion,ymin=`2.5%`, ymax=`97.5%`,group=discount.rate),
                       color = 'lightgray',size=1,
                       position = position_dodge(0.8))+
        geom_point(aes(color = discount.rate), position = position_dodge(0.8),size=2)+
        coord_flip()+
        scale_color_brewer('Discount Rate',palette = 'Spectral',direction=-1)+
        scale_x_discrete(labels = function(x) str_wrap(x,width=30))+
        scale_y_continuous(limits=c(0,3000), label = scales::dollar)+
        #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
        #scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
        #scale_size_manual(values=c(rep(1,5),1.5,1))+
        xlab("")+ylab("$(2020)/mT" ~CH[4])+
        ggtitle(paste0('Net Present Damages per metric ton methane \nby GBD Super Region'))+
        theme_minimal()+
        theme(legend.position = c(0.7,0.4),
              legend.title=element_text(size=16), 
              legend.text=element_text(size=16))+
        theme(axis.text = element_text(size=16),
              axis.title = element_text(size=16))#+
        #geom_text(data=means, 
        #    aes(y=means, x=SuperRegion, group=discount.rate, 
        #        label=paste0('$', round(means, 2)), vjust=-1.5),
        #    #position = position_dodge(1.3),
        #    color    = 'grey20',
        #    size     = 4) 
        #scale_x_continuous(breaks=seq(2020,2100,10))
        #scale_y_continuous(limits=c(0,11500))
p1

p2 <-  cResultslong %>%
        filter(discount.rate == '2.0% Ramsey') %>%
        ggplot(aes(x=SuperRegion, y=mean,fill = SuperRegion))+
        geom_bar(stat = 'identity', position = 'dodge',width=0.6)+
        geom_pointrange(aes(x=SuperRegion,ymin=`2.5%`, ymax=`97.5%`),
                        position = position_dodge(width=0.2),size=0.3,alpha=0.8)+#errobar,width = 0.4)+
        coord_flip()+
        scale_fill_brewer(palette = 'Spectral',direction=-1)+
        scale_x_discrete(labels = function(x) str_wrap(x,width=30))+
        scale_y_continuous(limits=c(0,3000), label = scales::dollar)+
        #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
        #scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
        #scale_size_manual(values=c(rep(1,5),1.5,1))+
        xlab("")+ylab("$(2020)/mT" ~CH[4])+
        ggtitle(paste0('Damages per metric ton methane \nby GBD Super Region'))+
        theme_minimal()+
        theme(legend.position = 'none')+#c(0.7,0.3))+
        theme(axis.text = element_text(size=16),
              axis.title = element_text(size=16))+
        geom_text(data=means%>%filter(discount.rate == '2.0% Ramsey'),
            aes(y=means, x=SuperRegion,
                label=paste0('$', round(means, 2)), vjust=-0.35,hjust = -0.35),
            #position = position_dodge(1.3),
            color    = 'grey20',
            size     = 6)
        #scale_x_continuous(breaks=seq(2020,2100,10))
        #scale_y_continuous(limits=c(0,11500))
p2

ggsave(plot = p1, filename = 
         file.path('Results','Global BenMAP WebTool','Plots','npd_country_bar_discount.png'),width = 10,height = 5)
ggsave(plot = p2, filename = 
         file.path('Results','Global BenMAP WebTool','Plots','npd_country_bar_Ram2.png'),width = 10,height = 5)

```

#
```{r globaldiscountbars}

#create bar graph of damages by region

Results <- file.path('Results')
globalresults <-
  read_csv(file.path(Results,"npd_global_damage_transformation_woutCessation.csv"),
           col_types = cols())

globalresults <- 
  globalresults  %>%
  group_by_at(.vars = c('Model','discount.rate')) %>%
  arrange(desc(mean)) %>%
  filter(Model =='MMM') %>%
  mutate(SuperRegion = 'Global')%>%
  filter(discount.rate %in% 
           c('1.5% Ramsey', '2.0% Ramsey', '2.5% Ramsey', "3.0% Ramsey",'2.0% CDR','3.0% CDR'))

## ordering of central rate for alpha transparency
globalresults$discount.rate <- factor(globalresults$discount.rate, 
                                  levels=c('3.0% CDR','2.0% CDR',
                                          "3.0% Ramsey","2.5% Ramsey", 
                                          "2.0% Ramsey", "1.5% Ramsey" ))

## function for labels
means = globalresults %>% 
  group_by(discount.rate) %>% 
  summarise(means=mean) %>% 
  mutate(across(c('means'), ~ signif(.,3)))

p1 <-  ggplot(globalresults, aes(x=discount.rate, y=mean))+
        geom_point(stat = 'identity')+
        geom_pointrange(aes(x=discount.rate,ymin=`2.5%`, ymax=`97.5%`),size=0.3,alpha=0.8)+#errobar,width = 0.4)+
        coord_flip()+
        scale_fill_brewer(palette = 'Spectral',direction=-1)+
        scale_x_discrete(labels = function(x) str_wrap(x,width=30))+
        scale_y_continuous(limits=c(500,3500), label = scales::dollar)+
        #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
        #scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
        #scale_size_manual(values=c(rep(1,5),1.5,1))+
        xlab("")+ylab("$(2020)/mT" ~CH[4])+
        ggtitle(paste0('Global NPD per metric ton methane'))+
        theme_minimal()+
        theme(legend.position = 'none')+
        theme(axis.text = element_text(size=16),
              axis.title = element_text(size=16))+
        geom_text(data=means, 
            aes(y=means, x=discount.rate, group=discount.rate, 
                label=paste0('$', round(means, 2)), vjust=-1.5),
            #position = position_dodge(1.3),
            color    = 'grey20',
            size     = 4) 
        #scale_x_continuous(breaks=seq(2020,2100,10))
        #scale_y_continuous(limits=c(0,11500))
p1

ggsave(plot = p1, filename = 
         file.path(Results,'Global BenMAP WebTool','Plots','Global_NPD_by_discounting.png'),width = 10,height = 5)

```

```{r tornado}

#create a tornado plot with npd results from:
#1. different GCMs (created from monetization code?)
#2. BenMAP (created from monetization code?)
#3. Disocunting (created from monetization code)
#3. BenMAP socioeconomic scenarios (from methane-ozone RFT)
#4. NOx (**)

#Read in 
Results <- file.path('Results')
globalresults <-
  read_csv(file.path(Results,"npd_global_damage_transformation_NoCessation.csv"),
           col_types = cols())
Results2 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
                      'npd_global_rff_means_MMM.csv')
rft_results <- read_csv(Results2, col_types = cols())
nox_results1 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
                      'npd','npd_global_rff_means_NOx_1_MMM.csv')
nox_results1 <- read_csv(nox_results1, col_types = cols())%>%
  mutate(NOx = '1')
nox_results5 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
                      'npd','npd_global_rff_means_NOx_0.5_MMM.csv')
nox_results5 <- read_csv(nox_results5, col_types = cols())%>%
  mutate(NOx = '0.5')
nox_results15 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
                      'npd','npd_global_rff_means_NOx_1.5_MMM.csv')
nox_results15 <- read_csv(nox_results15, col_types = cols()) %>%
  mutate(NOx = '1.5')
nox_total_results <- rbind(nox_results1,nox_results5,nox_results15)

#gcm_results <- globalresults %>% filter(discount.rate == '2.0% Ramsey')
#gcm_results$diff <- gcm_results$mean - gcm_results$mean[gcm_results$Model=='MMM']
#gcm_results_min <- gcm_results %>% slice_min(diff)
#gcm_results_max <- gcm_results %>% slice_max(diff)

#cols: Group, Sub-Group, Min, Max, Mean
group_col <- c(rep("CRF (95% CI)",2),rep("GCM (mean)",5),rep("Socioeconomics (95% CI)",2),
                          rep("Discount Rate (mean)",5),rep('NOx Emissions (\u0394 50%)',2))
subgroup_col <- (c("CRF Upper","CRF Lower","HadGEM","CESM2","MIROC","GISS","GFDL","SocioEcon. Upper","SocioEcon. Lower",
                  "1.5% Ramsey","2.5% Ramsey","3.0% Ramsey","2.0% CDR",'3.0% CDR','-50% NOx','+50% NOx'))
val_col <- c(globalresults$`97.5%`[globalresults$Model=='MMM' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$`2.5%`[globalresults$Model=='MMM' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='HadGEM' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='CESM2' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='MIROC' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='GISS' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='GFDL' & globalresults$discount.rate=='2.0% Ramsey'],
             rft_results$npd_97.5_wlag[rft_results$discount.rate=='2.0% Ramsey'],
             rft_results$npd_2.5_wlag[rft_results$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='1.5% Ramsey'],
             #globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='2.0% Ramsey'],
             globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='2.5% Ramsey'],
             globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='3.0% Ramsey'],
             globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='2.0% CDR'],
             globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='3.0% CDR'],
             nox_total_results$mean_npd[nox_total_results$NOx=='0.5' & nox_total_results$discount.rate=='2.0% Ramsey'],
             nox_total_results$mean_npd[nox_total_results$NOx=='1.5' & nox_total_results$discount.rate=='2.0% Ramsey']
             )

#calulate the absolute difference
val_col <- val_col - globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='2.0% Ramsey']

plot_data <- data.frame(group_col, subgroup_col, val_col)

p1 <- plot_data %>%
  mutate(subgroup_col = (fct_relevel(subgroup_col, '3.0% CDR',
                                     "2.0% CDR","3.0% Ramsey",
                                     "2.5% Ramsey","1.5% Ramsey",
                                     "MIROC","GISS","GFDL",
                                     "CESM2","HadGEM",
                                     '-50% NOx','+50% NOx',
                                     "SocioEcon. Lower","SocioEcon. Upper",
                                     "CRF Lower","CRF Upper"
                                     ))) %>%
  ggplot( aes(x=(subgroup_col), y=(val_col),fill=group_col))+
  geom_bar(stat='identity',width = 0.4)+
  #scale_fill_brewer() (palette = 'Spectral',direction=-1)+
  scale_fill_manual(name = '',values=c(brewer.pal(11,"Spectral")[c(2,4,8,10,11)]),
  breaks= c('CRF (95% CI)','Socioeconomics (95% CI)','NOx Emissions (\u0394 50%)','GCM (mean)','Discount Rate (mean)'))+
  coord_flip()+
  theme_minimal()+
  xlab('')+
  ylab("Change in Global $1660(2020)/mT" ~CH[4])+
  theme(axis.text = element_text(size=16),
        axis.title = element_text(size=16))+
  scale_y_continuous(limits = c(-1000, 1300),label = scales::dollar)+
  theme(legend.position = c(0.8,0.25),
        legend.title=element_text(size=14), 
        legend.text=element_text(size=14))

p1

ggsave(plot = p1, filename = 
         file.path(Results,'Global BenMAP WebTool','Plots','Global Uncertainty.png'),width = 10,height = 5)

```

### 2023-05-11 analysis/Manuscript/BenMAP/BenMap_O3Mort_Valuation.R

@@ -1,0 +1,426 @@
+#####
+## Title:   BenMAP_O3Mort_Valuation.R
+## Purpose: This file reads in O3 Mortality results from BenMAP
+##          and calculates the annual damages$ and 
+##          calculates the NPD value ($ per ton CO2) using 
+##          1.5%, 2%, 2.5%, and 3% Ramsey discounting and a 2% and 3% constant discount rate
+##          by 2100 in 2020 USD.
+## Inputs:  
+## Outputs: output/npd/co2_full_streams_2020_national_default_adaptation.parquet
+##          output/npd/npd_fredi_national.csv
+## Written by: US EPA, Climate Change Division (OAP); March 2023
+## Last updated: 3.30.23 by E. McDuffie
+## NOTES: 
+#####
+
+##########################
+#################  library
+##########################
+
+## clear workspace
+rm(list = ls())
+gc()
+
+## this function will check if a package is installed, and if not, install it
+list.of.packages = c('tidyverse', 'magrittr',
+                     'arrow',
+                     'zoo',
+                     'doParallel', 'foreach',
+                     'ggplot2')
+new.packages = list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages, repos = "http://cran.rstudio.com/")
+lapply(list.of.packages, library, character.only = TRUE)
+
+##########################
+################### constants
+##########################
+
+
+Inputs_mort <- file.path('analysis','Manuscript','BenMAP','Outputs')
+Inputs_socio <- file.path('input','RFF') #need mean population that was used in BenMAP, as well as the corresponding GDP scenario
+cntryxwalkpath <- file.path('input')
+Results <- file.path('analysis','Manuscript','BenMAP','Outputs')
+
+#### Constants
+base_vsl = 9.52e6 #us vsl in 2020$ 
+# EPA 2010 (updated 2014) guidance = $7.9 million in 2008$ (= 9.52mil in 2020$)
+Elasticity = 1
+
+#read in the O3 mortality data (interpolated)
[…]

### 2023-05-11 analysis/Manuscript/BenMAP/Inputs/2022 Global Methane Assessment/makeshape.py

@@ -1,0 +1,49 @@
+from shapely.geometry import box
+import geopandas as gpd
+import os
+import numpy as np
+from glob import glob
+import xarray as xr
+import cdo
+
+
+# Use Climate Data Operators to Calculate Area
+cdoo = cdo.Cdo('/work/ROMO/anaconda_envs/cdo-1.9.8/bin/cdo')
+
+# Choose a file at random to represent grid structure and cell area
+paths = sorted(glob('/work/MOD3DEV/kseltzer/GMA/metrics/*/O3_metrics_*_Sim?_05x05.nc'))
+gf = xr.open_dataset(paths[0])
+areaf = cdoo.gridarea(input=paths[0], returnCdf=True)
+AREA = areaf.variables['cell_area'][:]
+
+# Create a vector of boxes and areas
+boxes = []
+areas = []
+colvals = []
+rowvals = []
+i = 0
+n = AREA.size
+print('Building shapes')
+# Iterating column on the inner dimension. This seems consistent with received shapefile
+# I hope it doesn't really matter, but I don't know.
+for ri, lat in enumerate(gf.latitude):
+    for ci, lon in enumerate(gf.longitude):
+        i += 1
+        print(f'\r{i/n:.1%}', end='')
+        mybox = box(lon - 0.25, lat - 0.25, lon + 0.25, lat + 0.25)
+        boxes.append(mybox)
+        areas.append(AREA[ri, ci])
+        # using 1-base to be consistent with previous files
+        colvals.append(ci + 1)
+        # using 1-base to be consistent with previous files
+        rowvals.append(ri + 1)
+
+print('\nCreating vector object')
+gdf = gpd.GeoDataFrame(
+    dict(AREA=areas, ROW=rowvals, COL=colvals),
+    geometry=np.array(boxes),
+    crs=4326
+)
+
+print('Saving file')
+gdf.to_file("shp/GMA_05x05.shp")

### 2023-05-11 analysis/Manuscript/BenMAP/Inputs/2022 Global Methane Assessment/tobenmap.py

@@ -1,0 +1,58 @@
+import os
+import numpy as np
+from glob import glob
+import xarray as xr
+
+paths = sorted(glob('/work/MOD3DEV/kseltzer/GMA/metrics/*/O3_metrics_*_Sim?_05x05.nc'))
+outkeys = ['SUMMER_MDA8', 'ANNUAL_MDA8']
+
+for path in paths:
+    bpath = os.path.basename(path)[:-3]
+    print(bpath, end=':')
+    f = xr.open_dataset(path)
+    # Create place holders for outputs
+    f['SUMMER_MDA8'] = f['MDA8_NH'].copy() * 0
+    f['ANNUAL_MDA8'] = f['MDA8_NH'].copy() * 0
+    
+    # Overwrite SUMMER with only NH (where lat >= 0) or SH (where lat < 0)
+    f['SUMMER_MDA8'][:] = xr.concat(
+        [
+            f['MDA8_NH'].where(f['latitude'] >= 0),
+            f['MDA8_SH'].where(f['latitude'] < 0)
+        ], dim='season'
+    ).mean('season')
+    # Overwrite ANNUAL with the average of both seasons
+    f['ANNUAL_MDA8'][:] = xr.concat(
+        [f['MDA8_NH'], f['MDA8_SH']],
+        dim='season'
+    ).mean('season')
+    # Create a subset file
+    outf = f[outkeys]
+    nc = f.dims['longitude']
+    nr = f.dims['latitude']
+    # using 1-base to be consistent with previous files
+    col = np.arange(1, nc + 1)
+    row = np.arange(1, nr + 1)
+    COL, ROW = np.meshgrid(col, row)
+    outf['Col'] = xr.DataArray(COL, dims=('latitude', 'longitude'))
+    outf['Row'] = xr.DataArray(ROW, dims=('latitude', 'longitude'))
+
+    # Column is iterating first, then Row
+    # This is just for consistency with previous files. I don't think it matters
+    df = outf.transpose('longitude', 'latitude').to_dataframe()
+    
+    df['Metric'] = 'D8HOURMAX'
+    df['Seasonal Metric'] = ''
+    df['Statistic'] = ''
+    for outkey in outkeys:
+        print(outkey, end='.')
+        csvpath = os.path.join('csv', bpath + f'_{outkey}.csv')
+        # For each key, rename it as values and output it
+        df.rename(
+            columns={outkey: 'Values'}
+        ).loc[
[…]

### 2023-05-11 analysis/Manuscript/BenMAP/Inputs/Generate AQ Surfaces for BenMAP.R

@@ -1,0 +1,246 @@
+################################################################################
+## Title: Generate AQ Surfaces for BenMAP.R
+## Purpose: Process results from 5 global models (rom UNEP/CCAC Global Methane
+##          Assessment) and generate maps of O3 mixing ratios for input to BenMAP
+##          associated with the baseline (Sim #1) the baseline + ozone mixing
+##          ratios produced from a pulse of methane emissions. Perturbed O3 
+##          mixing ratios are calculated using the O3/CH4 response from the
+##          difference in O3 and CH4 in UNEP/CCAC Sims #1 and #2. Results
+##          are averaged across GCMs to calculate O3 from the MMM. All O3 maps
+##          are output for the full grid and for the subset of grid cells where
+##          there are people. Maps are calculated for methane concentrations
+##          at 11 time slices. The O3-CH4 response is held constant overtime.
+##    Inputs:
+##    UNEP/CCAC Global Methane Assessment Simulations for:
+##      -CESM2
+##      -GISS
+##      -GFDL
+##      -HadGEM (UK)
+##      -MIROC
+##    Pre-processed cross-walk of grid cells with population data
+##   Outputs:
+##    Full Globe & Population Subset maps for O3 mixing ratios from CCAC/UNEP Sim#1
+##    and the calculated ozone mixing ratio maps associated with the custom methane pulse
+##  Written by: Industrial Economics, Inc. Updated by EPA OAP
+##  Last Updated: May 2023
+####################################################################################
+
+#clear workspace
+rm(list = ls())
+
+#load in necessary packages
+  library(dplyr)
+  library(tidyr)
+  library(purrr)
+  library(patchwork)
+  library(arrow)
+
+#set working directory/paths
+  Inputs <- file.path(getwd(),'analysis','Manuscript', 'BenMAP','Inputs','2022 Global Methane Assessment','csv')
+  Outputs <- file.path(getwd(),'analysis','Manuscript', 'BenMAP','Inputs','AQ Surfaces')
+  
+  #set GCM & methane pulse variables
+  Pulse       = 275 #million metric tons
+  Models      <- c("CESM2","GFDL","GISS","HadGEM","MIROC") #5 UNEP/CCAC GCMs
[…]

### 2023-05-11 analysis/Manuscript/BenMAP/Summarizing BenMAP Results.R

@@ -1,0 +1,356 @@
+#################################################################################
+## Title: Summarizing BenMAP Results
+## Purpose:  This script unzips and combines the 6 models and 11 years of ozone mortality
+##    respiratory results from the BenMAP cloud application
+##    Output from BenMAP are dataframes of respiratory-related mortality at the
+##    country level, associated with the difference between O3 mixing ratios in
+##    the baseline scenario compared to O3 in the perturbed methane scenario. 
+## Inputs:
+##  - BenMAP cloud run output
+##  - COuntry name cross walk
+##  - 55km population counts
+## Ouputs:
+##  - csv files (interpolated and not) of annual respiratory-related mortality data
+##    associated with O3 from the marginal methane pulse, as well as the 
+##    total respiratory-mortality counts, BenMAP confidence intervals, and the
+##    population-weighted O3/CH4 response, globally and by country for each model
+##    and the calculated MMM
+## Written by: Industrial Economics, Inc., Updated by EPA OAP
+## last updated: May 2023
+## 
+#################################################################################
+rm(list = ls())
+
+# Install necessary libraries
+library(data.table)
+library(tidyverse)
+library(ggplot2)
+library(RColorBrewer)
+library(dplyr)
+library(scales)
+library(arrow)
+
+## 1. set directories
+  resultspath = file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')  
+  cntryxwalkpath = file.path('input')
+  aqspath = file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs','Aq Surfaces', 'Intermediate')
+  BenMapinputs = file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs')
+  #Plots <- file.path(getwd(), 'analysis','Manuscript','BenMAP','Ouptuts','Plots')
+
+## 2. Load BenMAP output data
+  # Creates a list of all zipped files in the input folder
+  zips<- dir(file.path(resultspath,'Cloud_runs'), pattern =".zip")
+
+  # Unzips all zipped files in the downloads folders
+#  walk(zips, ~ unzip(zipfile = str_c(paste0(resultspath,'/Cloud_runs/'), .x), 
+#                     overwrite = TRUE,
[…]

### 2023-05-11 analysis/Manuscript/Paper_RFigures.Rmd
Update to BenMAP valuation results

@@
 
+##Figure S5
+
+global maps of respiratory mortality, absolute total (by model)
 ```{r, countrymap}
 
-#should these be per capita?
-# add per capita code
-
-## FIgure 2b and adapted to S5
+## Figure 2b and adapted to S5
 
 #map of country mortality results
-MODEL = 'HadGEM'
-
-Plots <-file.path(getwd(),'Results','Global BenMAP WebTool')
-
-countryResults <- read.csv(paste0(Plots,"/Country Results Summary by Model & Year_eem.csv"))
+MODEL = 'MMM'
+
+Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
+input <- file.path(getwd(),'input','BenMAP')
+countryResults <- read.csv(paste0(input,"/Country Results Summary by Model & Year.csv"))
 
@@
 
-#plot by super region?
-
-#need to sum all country results
-
-#map data
+# load map data and align country names
 worldData <- map_data("world")
@@
                color = 'black',linewidth=0.15)+
-
-#  scale_fill_binned("Total Deaths",
-#                    breaks= c(0,100,1000,10000,80000),
-#                    guide= guide_colorsteps(even.steps = FALSE))+
   scale_fill_fermenter("Total Deaths",
@@
 
-# Map of the world:
-#p5<- map('world',col="grey", fill=TRUE, bg="white", lwd=0.05, mar=rep(0,4),border=0, ylim=c(-80,80) )
-
 p5
 ggsave(plot = p5, filename = 
-         file.path(Plots,'Plots',paste0('Total_MortMap_',MODEL,'.png')),width = 9,height = 5)
+         file.path(Plots,paste0('Total_MortMap_',MODEL,'.png')),width = 9,height = 5)
 ```
 
+## Figures 2b and SXX
+
+global maps of respiratory mortality, per capita (in 2020)
+as well as maps of 2020 population, mortality rate, and average O3/CH4 response
 ```{r, countrymap_percapita}
-
-#should these be per capita?
-# add per capita code
-
-## Figure 2b and adapted to S5
 
@@
 
-Plots <-file.path(getwd(),'Results','Global BenMAP WebTool')
-
-countryResults <- read.csv(paste0(Plots,"/Country Results Summary by Model & Year_eem.csv"))
+Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
+input <- file.path(getwd(),'input')
+countryResults <- read.csv(paste0(input,'/BenMAP',"/Country Results Summary by Model & Year.csv"))
+
 country_2020 <- countryResults %>%
@@
 #read in population
-inputs <- file.path("Code","InputData","Creating RFF Lookup Tables")
-mean.pop<-read.csv(file.path(inputs,"rff_pop_means.csv")) #this dataset includes all countries (rff-sp does not)
+socio_inputs <- file.path('input','RFF','lookup_tables')
+mean.pop<-read.csv(file.path(socio_inputs,"rff_pop_means.csv")) #this data set includes all countries (rff-sp does not)
 mean.pop <- mean.pop %>%
@@
 
-inputs3 <- file.path("Code","InputData","Tool Inputs")
-Ctry.Name<-read.csv(file.path(inputs3,'Final Country Grid Col_Row Index_EEM.csv'))
+#inputs3 <- file.path("Code","InputData","Tool Inputs")
+Ctry.Name<-read.csv(file.path(input,'Final Country Grid Col_Row Index.csv'))
 
@@
 
-#read in mort data (created in the mortality timeseries chunk)
+#read in mort data (created in the mortality timeseries chunk above)
 mort_rate <- resp_mort_means_noage %>% 
@@
 
-#plot by super region?
-
-#need to sum all country results
-
-#map data
+# read in map data and align country names
 worldData <- map_data("world")
@@
 
-#chloropleth of world
-# Percapita deaths
+#chloropleth of world Per capita deaths
 p5 <- ggplot() +
@@
                color = 'black',linewidth=0.15)+
-#  scale_fill_gradientn(
-#   colors=c("red","white","blue"),
-#   values=rescale(c(-1,0,1)),
-#   limits=c(-1,1))+
   scale_fill_fermenter("Deaths Per Capita",
@@
                                 0.3),
-                                #max(MergedData$death_percap, na.rm=TRUE)),
-                       #midpoint=1,
-                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
-                       #breaks= c(0,100,1000,10000,100000),
-                       #n.breaks = 6,
-                       #label=scales::comma)+#,
                        labels = c('Minimum',rep('',8),'Maximum'))+
-                       #guide =  guide_colorbar(nbin=10))+
-
-                       #guides() (fill= guide_colorbar(even.steps) = FALSE
-  #scale_fill_continuous(name="Total Mortality (deaths per 100K)",
-  #                      low="darkblue",
-  #                      high="lightblue",
-  #                      limits=c(0,70),
-  #                      na.value='gray')+
   theme_void()+
@@
         legend.title = element_text(size=13))+
-  #ggtitle('Total Deaths Per 100K')+
-  #theme(plot.title = element_text(size = 20))+
   guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
   coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
-
-# Map of the world:
-#p5<- map('world',col="grey", fill=TRUE, bg="white", lwd=0.05, mar=rep(0,4),border=0, ylim=c(-80,80) )
 

### 2023-05-11 analysis/Manuscript/BenMAP/BenMap_O3Mort_Valuation.R

@@ -1,17 +1,21 @@
-#####
+##################################################################
 ## Title:   BenMAP_O3Mort_Valuation.R
 ## Purpose: This file reads in O3 Mortality results from BenMAP
 ##          and calculates the annual damages$ and 
 ##          calculates the NPD value ($ per ton CO2) using 
 ##          1.5%, 2%, 2.5%, and 3% Ramsey discounting and a 2% and 3% constant discount rate
 ##          by 2100 in 2020 USD.
-## Inputs:  
-## Outputs: output/npd/co2_full_streams_2020_national_default_adaptation.parquet
-##          output/npd/npd_fredi_national.csv
+##          Option to implement cessation lags or not
+## Inputs:  Final Country Grid Col_Row Index.csv
+##          rffsp_pop_gdp_all_trials.parquet
+##          Country Results Interpolated 2020-2100 by Model.csv
+## Outputs: Country Results Interpolated Mortality Annual Undiscounted Damages_w[out]Cessation.csv
+##          npd_damage_transformation_full_streams_w[out]Cessation.parquet
+##          npd_global_damage_transformation_w[out]Cessation.csv
 ## Written by: US EPA, Climate Change Division (OAP); March 2023
-## Last updated: 3.30.23 by E. McDuffie
+## Last updated: 5.11.23 by E. McDuffie
 ## NOTES: 
-#####
+#####################################################################
 
 ##########################
 #################  library
@@ -42,85 +46,31 @@ cntryxwalkpath <- file.path('input')
 Results <- file.path('analysis','Manuscript','BenMAP','Outputs')
 
 #### Constants
-base_vsl = 9.52e6 #us vsl in 2020$ 
-# EPA 2010 (updated 2014) guidance = $7.9 million in 2008$ (= 9.52mil in 2020$)
+pulse = 275e6 #original methane pulse was 275 million metric tonnes
+base_vsl = 9.33e6 #us vsl in 2020$ 
+# from USEPA 2006 value in 2010 guidance 
 Elasticity = 1
+# turn 20-year cessation flag on (=1) or off (=0)
+cessation_flag = 1
 
[…]

### 2023-05-11 analysis/Manuscript/BenMAP/Summarizing BenMAP Results.R

@@ -15,6 +15,7 @@
 ##    total respiratory-mortality counts, BenMAP confidence intervals, and the
 ##    population-weighted O3/CH4 response, globally and by country for each model
 ##    and the calculated MMM
+##    Output data are used in the MOMM-RFT and are also saved in the RFT input/BenMAP folder
 ## Written by: Industrial Economics, Inc., Updated by EPA OAP
 ## last updated: May 2023
 ## 
@@ -31,9 +32,10 @@ library(scales)
 library(arrow)
 
 ## 1. set directories
-  resultspath = file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')  
+  resultspath = file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs') 
+  rft_inputpath = file.path(getwd(),'input','BenMAP')
   cntryxwalkpath = file.path('input')
-  aqspath = file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs','Aq Surfaces', 'Intermediate')
+  aqspath = file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs','AQ Surfaces', 'Intermediate')
   BenMapinputs = file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs')
   #Plots <- file.path(getwd(), 'analysis','Manuscript','BenMAP','Ouptuts','Plots')
 
@@ -345,12 +347,16 @@ library(arrow)
                                      by=c("column"="COL"))
   FormattedResults.Ctry<-FormattedResults.Ctry[,c(1,20,21,22,3,15,17:19,4:14)]
     
-  #export summary table
+  #export summary table (to two locations)
   write.csv(FormattedResults.Ctry,paste0(resultspath,"/Country Results Summary by Model & Year.csv"),row.names=F,na="")
+  write.csv(FormattedResults.Ctry,paste0(rft_inputpath,"/Country Results Summary by Model & Year.csv"),row.names=F,na="")
   
   #export interpolated results by country with ozone response
   write.csv(Interp.ByCtry,paste0(resultspath,"/Country Results Interpolated 2020-2100 by Model.csv"),
                               row.names=FALSE,na="")
+  write.csv(Interp.ByCtry,paste0(rft_inputpath,"/Country Results Interpolated 2020-2100 by Model.csv"),
+            row.names=FALSE,na="")
+  
 
 ## Have a nice day!    
     

### 2023-05-11 analysis/Manuscript/Paper_RFigures.Rmd
Updates to R figure files

@@
 knitr::opts_knit$set(root.dir = "~/CCD-CSIB/GitHub/Code/MOMM-RFT")
-#setwd("~/CCD-CSIB/GitHub/Code/MOMM-RFT")
 ```
@@
 library('zoo')
-#library('Rfast')
 ```
@@
 
-#Define BemMAP years and create dataframes
+#Define BenMAP years and create dataframes
 Model.years<-c(2020,2025,2030,2035,2040,2050,2060,2070,2080,2090,2100)
@@
 #Read in the globally averaged mean O3 response data
-# Generated by /analaysis/Manuscript/BenMAP/Inputs/Generate AQ Surfaces for BenMAP.R
+# Generated by /analysis/Manuscript/BenMAP/Inputs/Generate AQ Surfaces for BenMAP.R
 All.mean.results <- read.csv(file.path(Input,"AQ Surfaces","Mean_Pulse_Data.csv"))
@@
                        limits=c(0,100000),
-                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
                        breaks= c(0,100,1000,10000,100000),
-                       #n.breaks = 6,
-                       label=scales::comma)+#,
-                       #guide =  guide_colorbar(nbin=10))+
-
-                       #guides() (fill= guide_colorbar(even.steps) = FALSE
-  #scale_fill_continuous(name="Total Mortality (deaths per 100K)",
-  #                      low="darkblue",
-  #                      high="lightblue",
-  #                      limits=c(0,70),
-  #                      na.value='gray')+
+                       label=scales::comma)+
   theme_void()+
@@
         legend.title = element_text(size=13))+
-  #guide(c)
 
@@
                                 max(MergedData$MeanMortalityRate, na.rm=TRUE)),
-                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
-                       #breaks= c(0,100,1000,10000,100000),
-                       #n.breaks = 6,
-                       #label=scales::comma)+
                        labels = c('Minimum','','','','','','Maximum'))+
@@
                                 max(MergedData$AvgResponse, na.rm=TRUE)),
-                       #breaks= c(0,100,1000,5000,10000,20000,40000,60000,100000),
-                       #breaks= c(0,100,1000,10000,100000),
-                       #n.breaks = 6,
-                       #label=scales::comma)+
                        labels = c('Minimum','','','','','','Maximum'))+
@@
   theme(plot.title = element_text(size = 20))+
-  #guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
   coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
@@
 ggsave(plot = p, filename = 
-         file.path(Plots,'Plots',paste0('Relative_2020_CombMaps_',MODEL,'.png')),width = 9,height = 15)
+         file.path(Plots,paste0('Relative_2020_CombMaps_',MODEL,'.png')),width = 9,height = 15)
 
 ggsave(plot = p5, filename = 
-         file.path(Plots,'Plots',paste0('Relative_2020_DeathsPerCapMaps_',MODEL,'.png')),width = 9,height = 5)
-p5
+         file.path(Plots,paste0('Relative_2020_DeathsPerCapMaps_',MODEL,'.png')),width = 9,height = 5)
 ```
 
+## Figure 2c and SX
+
+barchart of NPD values by region, central case and by discount rate
 ```{r, damagesbar}
 
-#create bar graph of damages by region
-
-## FIgure 2c
-
+Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
+inputs <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
 countryresults <-
-  #read_csv(file.path('Results',"Country Results Interpolated Mortality Annual Undiscounted Damages.csv"),
-   #        col_types = cols())
-   read_csv(file.path('Results',"npd_country_damage_transformation_NoCessation.csv"),
+   read_csv(file.path(inputs,"npd_country_damage_transformation_woutCessation.csv"),
            col_types = cols())
@@
 globalresults <-
-  read_csv(file.path('Results',"npd_global_damage_transformation_NoCessation.csv"),
+  read_csv(file.path(inputs, "npd_global_damage_transformation_woutCessation.csv"),
            col_types = cols())
@@
   arrange(desc(mean)) %>%
-  #mutate(damages = damages/275e6,
-  #       damages_2_5 = damages_2_5/275e6,
-  #       damages_97_5 = damages_97_5/275e6) %>% #convert to per mt CH4 (not discounted)
   filter(Model =='MMM') %>%
-  #select(-c('TotalResults','TotalPct2_5','TotalPct97_5','Model')) %>%
   rbind(globalresults)
@@
 
-# cResultslong <- reshape2::melt(cResultslong, id="SuperRegion")
-# cResultslong$Model.years <- strtoi(str_replace(cResultslong$variable, 'X',''))
-# globalMMM <- globalResultslong[globalResultslong$Model == 'MMM - Mean AQS',]
-# globalMMM$SuperRegion <- 'Global'
-# globalMMM <- globalMMM %>%
-#   select(-c('Model'))
-# cResultslong <- rbind(cResultslong, globalMMM)
-
-#cResultslong <- cResultslong %>%
-#  group_by_at(c("COUNTRY","Model.years")) %>%
-#  arrange(desc(value)) %>%
-#  mutate(COUNTRY = factor(COUNTRY, levels=c(
-
   reg_i <- cResultslong %>% 
   group_by_at(.vars=c("SuperRegion")) %>% 
-  #filter(Model.years == 2020) %>%
   summarize_at(.vars=c("mean"),sum,na.rm=T) %>%
@@
   arrange_at(.vars=c("SuperRegion"))
+
+  # Figure SX
 
@@
         scale_y_continuous(limits=c(0,3000), label = scales::dollar)+
-        #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
-        #scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
-        #scale_size_manual(values=c(rep(1,5),1.5,1))+
         xlab("")+ylab("$(2020)/mT" ~CH[4])+
@@
         theme(axis.text = element_text(size=16),
-              axis.title = element_text(size=16))#+
-        #geom_text(data=means, 
-        #    aes(y=means, x=SuperRegion, group=discount.rate, 
-        #        label=paste0('$', round(means, 2)), vjust=-1.5),
-        #    #position = position_dodge(1.3),
-        #    color    = 'grey20',
-        #    size     = 4) 
-        #scale_x_continuous(breaks=seq(2020,2100,10))
-        #scale_y_continuous(limits=c(0,11500))
-p1
-
+              axis.title = element_text(size=16))
+# Figure 2c
 p2 <-  cResultslong %>%
@@
                 label=paste0('$', round(means, 2)), vjust=-0.35,hjust = -0.35),
-            #position = position_dodge(1.3),
             color    = 'grey20',
             size     = 6)
-        #scale_x_continuous(breaks=seq(2020,2100,10))
-        #scale_y_continuous(limits=c(0,11500))
 p2
@@
 ggsave(plot = p1, filename = 
-         file.path('Results','Global BenMAP WebTool','Plots','npd_country_bar_discount.png'),width = 10,height = 5)
+         file.path(Plots,'npd_country_bar_discount.png'),width = 10,height = 5)
 ggsave(plot = p2, filename = 
-         file.path('Results','Global BenMAP WebTool','Plots','npd_country_bar_Ram2.png'),width = 10,height = 5)
+         file.path(Plots,'Fig2c.png'),width = 10,height = 5)
 
@@
 
-#
+Extra
 ```{r globaldiscountbars}
 
-#create bar graph of damages by region
-
-Results <- file.path('Results')
-globalresults <-
-  read_csv(file.path(Results,"npd_global_damage_transformation_woutCessation.csv"),
-           col_types = cols())
-
-globalresults <- 
-  globalresults  %>%
-  group_by_at(.vars = c('Model','discount.rate')) %>%
-  arrange(desc(mean)) %>%
-  filter(Model =='MMM') %>%
-  mutate(SuperRegion = 'Global')%>%
-  filter(discount.rate %in% 
-           c('1.5% Ramsey', '2.0% Ramsey', '2.5% Ramsey', "3.0% Ramsey",'2.0% CDR','3.0% CDR'))
-
-## ordering of central rate for alpha transparency
-globalresults$discount.rate <- factor(globalresults$discount.rate, 
-                                  levels=c('3.0% CDR','2.0% CDR',
-                                          "3.0% Ramsey","2.5% Ramsey", 
-                                          "2.0% Ramsey", "1.5% Ramsey" ))
-
-## function for labels
-means = globalresults %>% 
-  group_by(discount.rate) %>% 
-  summarise(means=mean) %>% 
-  mutate(across(c('means'), ~ signif(.,3)))
-
-p1 <-  ggplot(globalresults, aes(x=discount.rate, y=mean))+
-        geom_point(stat = 'identity')+
-        geom_pointrange(aes(x=discount.rate,ymin=`2.5%`, ymax=`97.5%`),size=0.3,alpha=0.8)+#errobar,width = 0.4)+
-        coord_flip()+
-        scale_fill_brewer(palette = 'Spectral',direction=-1)+
-        scale_x_discrete(labels = function(x) str_wrap(x,width=30))+
-        scale_y_continuous(limits=c(500,3500), label = scales::dollar)+
-        #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
-        #scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
-        #scale_size_manual(values=c(rep(1,5),1.5,1))+
-        xlab("")+ylab("$(2020)/mT" ~CH[4])+
-        ggtitle(paste0('Global NPD per metric ton methane'))+
-        theme_minimal()+
-        theme(legend.position = 'none')+
-        theme(axis.text = element_text(size=16),
-              axis.title = element_text(size=16))+
-        geom_text(data=means, 
-            aes(y=means, x=discount.rate, group=discount.rate, 
-                label=paste0('$', round(means, 2)), vjust=-1.5),
-            #position = position_dodge(1.3),
-            color    = 'grey20',
-            size     = 4) 
-        #scale_x_continuous(breaks=seq(2020,2100,10))
-        #scale_y_continuous(limits=c(0,11500))
-p1
-
-ggsave(plot = p1, filename = 
-         file.path(Results,'Global BenMAP WebTool','Plots','Global_NPD_by_discounting.png'),width = 10,height = 5)
+# #create bar graph of damages by region
+# Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
+# inputs <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
+# countryresults <-
+#    read_csv(file.path(inputs,"npd_country_damage_transformation_woutCessation.csv"),
+#            col_types = cols())
+# 
+# globalresults <-
+#   read_csv(file.path(inputs, "npd_global_damage_transformation_woutCessation.csv"),
+#            col_types = cols())
+# 
+# Results <- file.path('Results')
+# globalresults <-
+#   read_csv(file.path(Results,"npd_global_damage_transformation_woutCessation.csv"),
+#            col_types = cols())
+# 
+# globalresults <- 
+#   globalresults  %>%
+#   group_by_at(.vars = c('Model','discount.rate')) %>%
+#   arrange(desc(mean)) %>%
+#   filter(Model =='MMM') %>%
+#   mutate(SuperRegion = 'Global')%>%
+#   filter(discount.rate %in% 
+#            c('1.5% Ramsey', '2.0% Ramsey', '2.5% Ramsey', "3.0% Ramsey",'2.0% CDR','3.0% CDR'))
+# 
+# ## ordering of central rate for alpha transparency
+# globalresults$discount.rate <- factor(globalresults$discount.rate, 
+#                                   levels=c('3.0% CDR','2.0% CDR',
+#                                           "3.0% Ramsey","2.5% Ramsey", 
+#                                           "2.0% Ramsey", "1.5% Ramsey" ))
+# 
+# 
+# ## function for labels
+# means = globalresults %>% 
+#   group_by(discount.rate) %>% 
+#   summarise(means=mean) %>% 
+#   mutate(across(c('means'), ~ signif(.,3)))
+# 
+# 
+# 
+# p1 <-  ggplot(globalresults, aes(x=discount.rate, y=mean))+
+#         geom_point(stat = 'identity')+
+#         geom_pointrange(aes(x=discount.rate,ymin=`2.5%`, ymax=`97.5%`),size=0.3,alpha=0.8)+#errobar,width = 0.4)+
+#         coord_flip()+
+#         scale_fill_brewer(palette = 'Spectral',direction=-1)+
+#         scale_x_discrete(labels = function(x) str_wrap(x,width=30))+
+#         scale_y_continuous(limits=c(500,3500), label = scales::dollar)+
+#         #scale_linetype_manual(values=c(rep("solid",9),"solid","solid"))+
+#         #scale_colour_manual(values=c("black",brewer.pal(7,"Spectral")))+
+#         #scale_size_manual(values=c(rep(1,5),1.5,1))+
+#         xlab("")+ylab("$(2020)/mT" ~CH[4])+
+#         ggtitle(paste0('Global NPD per metric ton methane'))+
+#         theme_minimal()+
+#         theme(legend.position = 'none')+
+#         theme(axis.text = element_text(size=16),
+#               axis.title = element_text(size=16))+
+#         geom_text(data=means, 
+#             aes(y=means, x=discount.rate, group=discount.rate, 
+#                 label=paste0('$', round(means, 2)), vjust=-1.5),
+#             #position = position_dodge(1.3),
+#             color    = 'grey20',
+#             size     = 4) 
+#         #scale_x_continuous(breaks=seq(2020,2100,10))
+#         #scale_y_continuous(limits=c(0,11500))
+# p1
+# 
+# ggsave(plot = p1, filename = 
+#          file.path(Results,'Global BenMAP WebTool','Plots','Global_NPD_by_discounting.png'),width = 10,height = 5)
 
 ```
-
+## Figure 3
+
+Tornado sensitivity plot
 ```{r tornado}
@@
 #2. BenMAP (created from monetization code?)
-#3. Disocunting (created from monetization code)
+#3. Discounting (created from monetization code)
 #3. BenMAP socioeconomic scenarios (from methane-ozone RFT)
@@
 #Read in 
+Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
+inputs <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
+rft_inputs <- file.path(getwd(),'output','npd')
+
 Results <- file.path('Results')
 globalresults <-
-  read_csv(file.path(Results,"npd_global_damage_transformation_NoCessation.csv"),
+  read_csv(file.path(inputs,"npd_global_damage_transformation_woutCessation.csv"),
            col_types = cols())
-Results2 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
-                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
-                      'npd_global_rff_means_MMM.csv')
+Results2 <- file.path(rft_inputs,'npd_global_rff_means_NOx_1_MMM.csv')
 rft_results <- read_csv(Results2, col_types = cols())
-nox_results1 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
-                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
-                      'npd','npd_global_rff_means_NOx_1_MMM.csv')
+nox_results1 <- file.path(rft_inputs,'npd_global_rff_means_NOx_1_MMM.csv')
 nox_results1 <- read_csv(nox_results1, col_types = cols())%>%
   mutate(NOx = '1')
-nox_results5 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
-                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
-                      'npd','npd_global_rff_means_NOx_0.5_MMM.csv')
+nox_results5 <- file.path(rft_inputs, 'npd_global_rff_means_NOx_0.5_MMM.csv')
 nox_results5 <- read_csv(nox_results5, col_types = cols())%>%
   mutate(NOx = '0.5')
-nox_results15 <- file.path('C:/Users/emcduf01','OneDrive - Environmental Protection Agency (EPA)',
-                      'Documents','CCD-CSIB','GitHub','Code','MOMM-RFT','output',
-                      'npd','npd_global_rff_means_NOx_1.5_MMM.csv')
+nox_results15 <- file.path(rft_inputs, 'npd_global_rff_means_NOx_1.5_MMM.csv')
 nox_results15 <- read_csv(nox_results15, col_types = cols()) %>%
@@
 nox_total_results <- rbind(nox_results1,nox_results5,nox_results15)
-
-#gcm_results <- globalresults %>% filter(discount.rate == '2.0% Ramsey')
-#gcm_results$diff <- gcm_results$mean - gcm_results$mean[gcm_results$Model=='MMM']
-#gcm_results_min <- gcm_results %>% slice_min(diff)
-#gcm_results_max <- gcm_results %>% slice_max(diff)
 
@@
 
-#calulate the absolute difference
+#calculate the absolute difference
 val_col <- val_col - globalresults$mean[globalresults$Model=='MMM' & globalresults$discount.rate=='2.0% Ramsey']
@@
 ggsave(plot = p1, filename = 
-         file.path(Results,'Global BenMAP WebTool','Plots','Global Uncertainty.png'),width = 10,height = 5)
+         file.path(Plots,'Fig3.png'),width = 10,height = 5)
 

### 2023-05-11 code/1_build_inputs_for_momm_rft.R
Update to RFT inputs, code, and results

@@ -35,7 +35,7 @@ options(scipen = 9999999)
 ##########################
 ##############  data paths
 ##########################
-setwd("~/shared/OAR/OAP/CCD/CSIB/Methane-Ozone/MOMM-RFT")
+#setwd("~/shared/OAR/OAP/CCD/CSIB/Methane-Ozone/MOMM-RFT")
 inputpath = file.path('input','RFF','pop_income')
 momm_rft_input_path = file.path('input','RFF','rft_inputs')
 
@@ -57,7 +57,7 @@ c_iteration  <- inputpath %>% list.files(pattern = "\\.feather") %>%
   (function(x){sub("rffsp_pop_income_run_", "", x)}) %>%
   as.numeric %>% sort; 
 c_iteration %>% length
-#c_iteration = c(1,2)
+c_iteration = c(1,2)
 
 
 rffsp <- 
@@ -97,4 +97,35 @@ rffsp <-
     write_parquet(momm_rft_input_path %>% file.path("rffsp_pop_gdp_all_trials.parquet"))
   
 
+### Process NOx Baseline Timeseries and NOx sensitivity equations
+  # (for each model)
+  
+models = c('CESM2','GFDL','GISS','HadGEM','MIROC','MMM')
+Results<-list()
+
+for (imodel in 1:length(models)) {
+  CCAC_temp <- read_excel(file.path('input','CH4_O3_v4_resp.xlsx'),
+                        sheet = paste0(models[imodel],"_results"))
+  print(models[imodel])
+  if (models[imodel] == 'MMM'){
+    CCAC_temp <- CCAC_temp[-1,c(1,3,16,19)]
+  } else {
+    CCAC_temp <- CCAC_temp[-1,c(1,3,34,37)]
+  }
+  colnames(CCAC_temp) = c('Country',"NOx_Mt","slope","intercept")
+  CCAC_temp$Model = models[imodel]
+  
+  Results[[imodel]] <- CCAC_temp
+  
+}
+All.results <- bind_rows(Results)
+
+#now join with country crosswalk and fill in missing countries
+countries <- read.csv(file.path('input',"Final Country Grid Col_Row Index_EEM.csv"))[,c(1,3:4,6:9)]
+
+Results2 <- right_join(All.results, countries, by = c("Country"="CCAC_country"),multiple='all')
+
+Results2 %>% 
+  write_csv(file.path("input","CCAC","Country_ozone_response_nox_sensitivity.csv"))
[…]

### 2023-05-11 code/2_Global Ozone Reduce Form Tool_EEM.R

@@ -1,31 +1,24 @@
-#### Purpose: To adjust the Global Ozone health effects impact generated in BenCloud####
+### Title: 2_Global Ozone Reduce Form Tool_EEM.R                                     ##
+### Purpose: To adjust the Global Ozone health effects impact generated in BenCloud  ####
 ### Created by: Melanie Jackson, IEc 
 ### Adapted by: E. McDuffie (EPA)
 ### Date Created: 3/15/2023      
-### Last Edited: 4/12/2023
+### Last Edited: 4/24/2023
 ##      Inputs:                                                                       ##
-##        -_Simulation Inputs.xlsx                                                    ##
 ##        -Country Results Interpolated 2020-2100 by Model.csv                        ##
 ##        -All Trajectory Mortality Ratios_through 2100.rds (parts 1&2)               ##
 ##        -All Trajectory Population Ratios_through 2100.rds (parts 1&2)              ##
-##        -RFF Country GDPperCap_interpolated through 2100.rds (23 parts)             ##
-##        -RFF Global GDP Per Capita_through 2100.csv                                 ##
-##        -Global 2100 Population Rank Order.xlsx                                     ##
 ##        -sampled_pop_trajectory_numbers.csv                                         ##
-##        -Discount Rates w_20yr Cessation Lags.xlsx                                  ##
-##        -Country Ozone Response.csv                                                 ##
-##        -BenMAP Inflation Indexed to 2015.xlsx                                      ##
-##        -Discount Rates w_20yr Cessation Lags.xlsx                                  ##
 ##        -Final Country Grid Col_Row Index.csv                                       ##
-##        -Ramsey Discounting.xlsx                                                    ##
 ##     Outputs:
 ##        - output/rft/
-## Units: Code adjusts all $ values for $2020, Ramsey discounted $ is PV to 2020      ##
+## Units: Code adjusts all $ values for $2020                                         ##
[…]

### 2023-05-11 code/3_methane_ozone_npd.R

@@ -1,23 +1,32 @@
-#####
-## Title:   3_methane_ozone_npd.R
-## Purpose: This file will read all output files from the Methane-Ozone Mortality
-##          Model-Reduced Form Tool and calculate net present damages
-##          and aggregate statistics across all trials.
-##          
-## Inputs:  output/damages/rffsp/damages_[scenario].parquet
-## Outputs: output/damages/rffsp/damages_[scenario].parquet
-## Written by: US EPA, Climate Change Division (OAP); April 2023
-## Last updated: 4/12/2023 by E. McDuffie
-## Notes:       
-##
-#####
+#######################################################################################
+### Title:        3_methane_ozone_npd.R
+### Purpose:      This file will read all output files from the Methane-Ozone Mortality
+##                Model-Reduced Form Tool and calculate net present damages
+##                and aggregate statistics across all trials. (for each model & country)
+##                The calculations are done for each model and country and combined at
+##                the end to help reduce the memory usage when processing all 10,000 RFF scenarios
+### Written by:   US EPA, Climate Change Division (OAP), adapted from NCEE (OP)
+### Date Created: 3/15/2023      
+### Last Edited:  4/25/2023
+##      Inputs:                                                                       ##
+##        -damages_mean_NOx_X_momm_rft.parquet                                        ##
+##        -Final Country Grid Col_Row Index.csv                                       ##
+##     Outputs:
+##        -npd_full_streams_rff_NOx_X_MODEL_COUNTRY.parquet                           ##
+##        - npd_country_rff_means_NOX_MODEL_COUNTRY.parquet                           ##
+##        - npd_global_rff_means_NOX_MODEL.csv                                        ##
+## Units: All $ values for $2020                                                      ##
+########################################################################################
+
 rm(list = ls()); gc()
 #necessary packages
+
[…]

### 2023-05-11 input/RFF/lookup_tables/Create RFF lookup Tables_eem.R

@@ -1,0 +1,385 @@
+#### Purpose: To create a look-up table of the RFF scenarios              ####
+##      Look-up table will contain the ratio of mortality/population      ##
+##          for each scenario to the mean mortality/population values     ##
+##      Ratios will be used to scale to the mean scenario point estimates ##
+############################################################################
+
+#### Load necessary packages ####
+  library(dplyr)
+  library(data.table)
+  library(readxl)
+  library(tidyverse)
+  library(zoo)
+  library(arrow)
+  library(progress)
+
+#### Set working directory and read in inputs ####
+  #Working Directories
+    inpath<-"J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Data/rff_data/new_dat/"
+    outpath<-"J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Results/Reduced Form Tool/RFF Lookup Tables/"
+    input <- file.path('Code','InputData','Creating RFF Lookup Tables')
+    input2 <- file.path('Code','InputData','RFF Lookup Tables')
+    
+  #input data (takes a few minutes to load, these files are large)
+    mx<-read.csv(paste0(input,"/mx_trajectories.csv"))
+    pop<-read.csv(paste0(input,"/pop_trajectories.csv"))
+    #subset for years <=2100
+      #must crosswalk mortality year ranges to years first
+        Cross.year<-read_xlsx("J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Data/Mortality/Crosswalk RFF Mort Year.xlsx")
+        mx<-left_join(mx,Cross.year,by=c("Year"="RFF year"))
+      mx<-mx[(mx$Min.Year<2100),]
+      pop<-pop[(pop$Year<=2100),]
+      #export files as rds so easier to load in other instances
+        saveRDS(mx,file=paste0(outpath,"All Mortality Data_through 2100.rds"))
+        saveRDS(pop,file=paste0(outpath,"All Population Data_through 2100.rds"))
+  #IFs resp adjustments - requires crosswalking IFs country ID to RFF country ID
[…]

### 2023-05-11 analysis/Manuscript/Paper_RFigures.Rmd
Added pop/mort pre-processing scripts & data

@@
 #IF Data on ratios of respiratory to total mortality rates
-IFsFactors<-read.csv(file.path(inputs,'IFP',"adjustment_factors.csv"))  
-IFsCountry<-read.csv(file.path(inputs,'IFP',"country_codes.csv"))
+IFsFactors<-read.csv(file.path(inputs3,'IFP',"adjustment_factors.csv"))  
+IFsCountry<-read.csv(file.path(inputs3,'IFP',"country_codes.csv"))
   names(IFsCountry)[1] <- 'RegionName2'
-IFsCtryCross<-read.csv(file.path(inputs,'IFP',"country_crosswalk.csv"))
+IFsCtryCross<-read.csv(file.path(inputs3,'IFP',"country_crosswalk.csv"))
 IFsFactors<-left_join(IFsFactors,IFsCountry,by=c("Region"="RegionNum"))
@@
 IFsFactors<-IFsFactors[,2:8]
-Cross.year<-read_xlsx(file.path(inputs,'IFP','Crosswalk RFF Mort Year.xlsx'))
+Cross.year<-read_xlsx(file.path(inputs3,'IFP','Crosswalk RFF Mort Year.xlsx'))
 mean.pop<-read.csv(file.path(inputs,'RFF',"rff_pop_means.csv"))

### 2023-05-11 code/2_Global Ozone Reduce Form Tool.R

@@ -31,11 +31,9 @@
     library(foreach) #to do parallel processing in loop
     library(purrr)  #for general data manipulation
     library(utils)
-    #ibrary(pbmcapply) #adds ETA to progress bar
     library(arrow)
   
   #working directories
-    #setwd("~/shared/OAR/OAP/CCD/CSIB/Methane-Ozone/MOMM-RFT")
     Inputs  <- file.path('input')
     Outputs <- file.path('output','rft')
     

### 2023-05-11 code/BenMAP/BenMap_O3Mort_Valuation.R

@@ -1,457 +1,0 @@
-#####
-## Title:   BenMAP_O3Mort_Valuation.R
-## Purpose: This file reads in O3 Mortality results from BenMAP
-##          and calculates the annual damages$ and 
-##          calculates the NPD value ($ per ton CO2) using 
-##          1.5%, 2%, 2.5%, and 3% Ramsey discounting and a 2% and 3% constant discount rate
-##          by 2100 in 2020 USD.
-## Inputs:  
-## Outputs: output/npd/co2_full_streams_2020_national_default_adaptation.parquet
-##          output/npd/npd_fredi_national.csv
-## Written by: US EPA, Climate Change Division (OAP); March 2023
-## Last updated: 3.30.23 by E. McDuffie
-## NOTES: 
-#####
-
-##########################
-#################  library
-##########################
-
-## clear workspace
-rm(list = ls())
-gc()
-
-## this function will check if a package is installed, and if not, install it
-list.of.packages = c('tidyverse', 'magrittr',
-                     'arrow',
-                     'zoo',
-                     'doParallel', 'foreach',
-                     'ggplot2')
-new.packages = list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-if(length(new.packages)) install.packages(new.packages, repos = "http://cran.rstudio.com/")
-lapply(list.of.packages, library, character.only = TRUE)
-
-##########################
-################### constants
-##########################
-
-
-Inputs_mort <- file.path('Results','Global BenMAP WebTool')
-Inputs_socio <- file.path('Code','InputData','Tool Inputs','RFF') #need mean population that was used in BenMAP, as well as the corresponding GDP scenario
-cntryxwalkpath <- file.path('Code/InputData/Tool Inputs')
-Results <- file.path('Results')
-
-#### Constants
-base_vsl = 9.52e6 #us vsl in 2020$ 
-## WHAT VSL TO USE?? ##
-# EPA 2010 (updated 2014) guidance = $7.9 million in 2008$ (= 9.52mil in 2020$)
-# What is used in FrEDI?
[…]

### 2023-05-11 input/RFF/lookup_tables/Create RFF lookup Tables.R

@@ -1,0 +1,385 @@
+##########################################################################
+## Title: Create RFF look-up Tables.R
+## Purpose: To create a look-up table of the RFF scenarios              ####
+##      Look-up tables will contain the ratio of (1)mortality and (2) population
+##          for each scenario to the mean (1) mortality and (2) population values
+##      Ratios will be used as input to the Reduced Form Tool ##
+## Written by: Industrial Economics, Inc. Edited by: EPA OAP
+## Last updated: May, 2023
+############################################################################
+
+#### Load necessary packages ####
+  library(dplyr)
+  library(data.table)
+  library(readxl)
+  library(tidyverse)
+  library(zoo)
+  library(arrow)
+  library(progress)
+
+#### Set working directory and read in inputs ####
+  #Working Directories
+    #inpath<-"J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Data/rff_data/new_dat/"
+    #outpath<-"J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Results/Reduced Form Tool/RFF Lookup Tables/"
+    input_rff   <- file.path('input','RFF','lookup_tables','input_data')
+    input_ifp   <- file.path('input','IFP') 
+    #input2  <- file.path('Code','InputData','RFF Lookup Tables')
+    outpath <- file.path('input','RFF','lookup_tables')
+    
+  #input data (takes a few minutes to load, these files are large)
+    mx  <- read.csv(paste0(input_rff,"/mx_trajectories.csv"))
+    pop <- read.csv(paste0(input_rff,"/pop_trajectories.csv"))
+    #subset for years <=2100
+    #must crosswalk mortality year ranges to years first
+    Cross.year <- read_xlsx(file.path(input_ifp,"Crosswalk RFF Mort Year.xlsx"))
+    mx  <- left_join(mx,Cross.year,by=c("Year"="RFF year"))
+    mx  <- mx[(mx$Min.Year<2100),]
+    pop <- pop[(pop$Year<=2100),]
+    #export files as rds so easier to load in other instances
+    saveRDS(mx,file=paste0(outpath,"All Mortality Data_through 2100.rds"))
[…]

### 2023-05-11 input/RFF/lookup_tables/Create RFF lookup Tables_eem.R

@@ -1,385 +1,0 @@
-#### Purpose: To create a look-up table of the RFF scenarios              ####
-##      Look-up table will contain the ratio of mortality/population      ##
-##          for each scenario to the mean mortality/population values     ##
-##      Ratios will be used to scale to the mean scenario point estimates ##
-############################################################################
-
-#### Load necessary packages ####
-  library(dplyr)
-  library(data.table)
-  library(readxl)
-  library(tidyverse)
-  library(zoo)
-  library(arrow)
-  library(progress)
-
-#### Set working directory and read in inputs ####
-  #Working Directories
-    inpath<-"J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Data/rff_data/new_dat/"
-    outpath<-"J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Results/Reduced Form Tool/RFF Lookup Tables/"
-    input <- file.path('Code','InputData','Creating RFF Lookup Tables')
-    input2 <- file.path('Code','InputData','RFF Lookup Tables')
-    
-  #input data (takes a few minutes to load, these files are large)
-    mx<-read.csv(paste0(input,"/mx_trajectories.csv"))
-    pop<-read.csv(paste0(input,"/pop_trajectories.csv"))
-    #subset for years <=2100
-      #must crosswalk mortality year ranges to years first
-        Cross.year<-read_xlsx("J:/share/OAQPS 2021/BenMAP FollowOn (TO 6)/Task 2.2 BenMAP Cloud Tool/2.2.4 Climate and International Analyses/Data/Mortality/Crosswalk RFF Mort Year.xlsx")
-        mx<-left_join(mx,Cross.year,by=c("Year"="RFF year"))
-      mx<-mx[(mx$Min.Year<2100),]
-      pop<-pop[(pop$Year<=2100),]
-      #export files as rds so easier to load in other instances
-        saveRDS(mx,file=paste0(outpath,"All Mortality Data_through 2100.rds"))
-        saveRDS(pop,file=paste0(outpath,"All Population Data_through 2100.rds"))
-  #IFs resp adjustments - requires crosswalking IFs country ID to RFF country ID
[…]

### 2023-05-12 analysis/Manuscript/Paper_RFigures.Rmd
Update to R figures script

@@
 ```
+
+## Figure S2
+Made elsewhere in python
 
@@
 ggsave(plot = p1, filename = 
-         file.path(Plots,'Timeseries_global_population.png'),width = 8,height = 5)
+         file.path(Plots,'FigS3a.png'),width = 8,height = 5)
 
@@
 ggsave(plot = p2, filename = 
-         file.path(Plots,'Timeseries_global_avgMortRate.png'),width = 8,height = 5)
+         file.path(Plots,'FigS3b.png'),width = 8,height = 5)
 
@@
 #plot line graph of global total and top 10 countries (for MMM)
-Plots <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
+Plots <- file.path(getwd(),'analysis','Manuscript','Plots')
+input <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
 
@@
 # from /analysis/Manuscript/BenMAP/Summarizing BenMAP Results.R
-countryResults    <- read.csv(paste0(Plots,"/Country Results Summary by Model & Year.csv"))
-globalResults     <- read.csv(paste0(Plots,"/Global Results Summary by Model & Year.csv"))
-globalResults2.5  <- read.csv(paste0(Plots,"/Global Results Summary by Model & Year 2_5CI.csv"))
-globalResults97.5 <- read.csv(paste0(Plots,"/Global Results Summary by Model & Year 97_5CI.csv"))
+countryResults    <- read.csv(paste0(input,"/Country Results Summary by Model & Year.csv"))
+globalResults     <- read.csv(paste0(input,"/Global Results Summary by Model & Year.csv"))
+globalResults2.5  <- read.csv(paste0(input,"/Global Results Summary by Model & Year 2_5CI.csv"))
+globalResults97.5 <- read.csv(paste0(input,"/Global Results Summary by Model & Year 97_5CI.csv"))
 
@@
 ggsave(plot = p1, filename = 
-         file.path(Plots,'Plots/Timeseries_O3Mort_GCMs.png'),width = 8,height = 5)
+         file.path(Plots,'FigS4.png'),width = 8,height = 5)
 
@@
 ggsave(plot = p5, filename = 
-         file.path(Plots,paste0('Total_MortMap_',MODEL,'.png')),width = 9,height = 5)
+         file.path(Plots,paste0('FigS5_',MODEL,'.png')),width = 9,height = 5)
 ```
 
-## Figures 2b and SXX
+## Figures 2b and S6
 
@@
 ggsave(plot = p, filename = 
-         file.path(Plots,paste0('Relative_2020_CombMaps_',MODEL,'.png')),width = 9,height = 15)
+         file.path(Plots,paste0('FigS6_',MODEL,'.png')),width = 9,height = 15)
 
 ggsave(plot = p5, filename = 
-         file.path(Plots,paste0('Relative_2020_DeathsPerCapMaps_',MODEL,'.png')),width = 9,height = 5)
+         file.path(Plots,paste0('Fig2b_',MODEL,'.png')),width = 9,height = 5)
 ```
 
-## Figure 2c and SX
+## Figure 2c and S7
 
@@
 ggsave(plot = p1, filename = 
-         file.path(Plots,'npd_country_bar_discount.png'),width = 10,height = 5)
+         file.path(Plots,'FigS7.png'),width = 10,height = 5)
 ggsave(plot = p2, filename = 

### 2023-05-12 analysis/Manuscript/Paper_RFigures.Rmd
Minor updates to rft output and figure script

@@
 rft_results <- read_csv(Results2, col_types = cols())
-nox_results1 <- file.path(rft_inputs,'npd_global_rff_means_NOx_1_MMM.csv')
+nox_results1 <- file.path(rft_inputs,'npd_global_means_NOx_1_MMM.csv')
 nox_results1 <- read_csv(nox_results1, col_types = cols())%>%
   mutate(NOx = '1')
-nox_results5 <- file.path(rft_inputs, 'npd_global_rff_means_NOx_0.5_MMM.csv')
+nox_results5 <- file.path(rft_inputs, 'npd_global_means_NOx_0.5_MMM.csv')
 nox_results5 <- read_csv(nox_results5, col_types = cols())%>%
   mutate(NOx = '0.5')
-nox_results15 <- file.path(rft_inputs, 'npd_global_rff_means_NOx_1.5_MMM.csv')
+nox_results15 <- file.path(rft_inputs, 'npd_global_means_NOx_1.5_MMM.csv')
 nox_results15 <- read_csv(nox_results15, col_types = cols()) %>%

### 2023-05-12 code/3_methane_ozone_npd.R

@@ -41,8 +41,9 @@ Outputs <- file.path("output",'npd')
 
 YEAR  = 2020
 PULSE = 275e6 #CH4 pulse of 275 million metric tonnes
+allRFF = 1
 
-countries <- read.csv(file.path('input',"Final Country Grid Col_Row Index_EEM.csv"))[,c(1,3:4,6,8)]
+countries <- read.csv(file.path('input',"Final Country Grid Col_Row Index.csv"))[,c(1,3:4,6,8)]
 countries <- countries %>% filter(Region != "")
 
 NOx_scalar = 1
@@ -68,15 +69,16 @@ read_momm_rft =
 
 
 for (MODEL in c('MMM')) { #},'CESM2','HadGEM','GISS','GFDL','MIROC')) {
-  # To read all files:
-  #damages = 
-  #  list.files(Inputs, full.names = T) %>% 
-  #  map_df(~read_momm_rft(., MODEL))
-
-  # to read specific file:
-  damages =
-    read_momm_rft(file.path(Inputs,paste0('damages_mean_NOx_',NOx_scalar,'_momm_rft.parquet')),MODEL)
-  
+  if (allRFF==1){
+    # To read all files:
+    damages = 
+      list.files(Inputs, full.names = T) %>% 
+      map_df(~read_momm_rft(., MODEL))
+  } else {
+    # to read specific file:
+    damages =
+      read_momm_rft(file.path(Inputs,paste0('damages_mean_NOx_',NOx_scalar,'_momm_rft.parquet')),MODEL)
+  }
   print(MODEL)
   
   countries <- damages %>% distinct(LocID)
@@ -154,9 +156,13 @@ for (MODEL in c('MMM')) { #},'CESM2','HadGEM','GISS','GFDL','MIROC')) {
   
 
   ## export full streams
- data %>%
+ if (allRFF ==1){
+  data %>%
     write_parquet(file.path(Outputs, paste0('npd_full_streams_rff_NOx_',NOx_scalar,'_',MODEL,'_',countries$LocID[COUNTRY],'.parquet')))
-#     write_parquet(file.path(Outputs, paste0('npd_full_streams_rff_',MODEL,'.parquet')))
+ } else {
+   data %>%
+     write_parquet(file.path(Outputs, paste0('npd_full_streams_NOx_',NOx_scalar,'_',MODEL,'_',countries$LocID[COUNTRY],'.parquet')))
+ }
 }
[…]

### 2023-05-23 analysis/Manuscript/Paper_RFigures.Rmd
Manuscript figure updates prior to submission

@@
 library('RColorBrewer')
-library('maps')
-library('mapproj')
+#library('maps')
+#library('mapproj')
 library('arrow')
@@
 library('zoo')
+library(sf)
+library(rnaturalearth)
+library(rnaturalearthdata)
 ```
@@
 #specify input and output paths
-Plots <-file.path(getwd(),'analysis','Manuscript','Plots')
-Input <- file.path(getwd(),'analysis','Manuscript','BenMAP','Inputs')
+dir.root <- getwd()
+Plots <-file.path('../..','analysis','Manuscript','Plots')
+Input <- file.path('../..','analysis','Manuscript','BenMAP','Inputs')
 
@@
 # Pre-processed in /code/1_build_inputs_for_momm_rft.R
-rffsp <- read_parquet(file.path(Inputs_socio, 'rffsp_pop_gdp_all_trials.parquet'))
+rffsp <- read_parquet(file.path(getwd(),Inputs_socio, 'rffsp_pop_gdp_all_trials.parquet'))
 
@@
 #plot line graph of global total and top 10 countries (for MMM)
-Plots <- file.path(getwd(),'analysis','Manuscript','Plots')
-input <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
+Plots <- file.path('../..','analysis','Manuscript','Plots')
+input <- file.path('../..','analysis','Manuscript','BenMAP','Outputs')
 
@@
 #map of country mortality results
-MODEL = 'MMM'
+MODEL = 'GFDL'
 
@@
   rename(region = COUNTRY) #%>%
-
-# load map data and align country names
-worldData <- map_data("world")
-#correct region names
-worldData["region"][worldData["region"] == "Russia"] <- "Russian Federation"
-worldData["region"][worldData["region"] == "USA"] <- "United States"
-worldData["region"][worldData["region"] == "Virgin Islands" & worldData['subregion']==' US'] <- "US Virgin Islands"
-worldData["region"][worldData["region"] == "Taiwan"] <- "China, Taiwan Province of China"
-worldData["region"][worldData["region"] == "Trinidad" | worldData['region']=='Tobago'] <- "Trinidad and Tobago"
-worldData["region"][worldData["region"] == "Grenadines"| worldData['region']=='Saint Vincent'] <- "Saint Vincent and the Grenadines"
-worldData["region"][worldData["region"] == "Swaziland"] <- "Eswatini"
-worldData["region"][worldData["region"] == "UK"] <- "United Kingdom"
-worldData["region"][worldData["region"] == "Cape Verde"] <- "Cabo Verde"
-worldData["region"][worldData["region"] == "Republic of Congo"] <- "Congo"
-worldData["region"][worldData["region"] == "Democratic Republic of the Congo"] <- "Congo DRC"
-worldData["region"][worldData["region"] == "Ivory Coast"] <- "Cote d'Ivoire"
-worldData["region"][worldData["region"] == "Brunei"] <- "Brunei Darussalam"
-worldData["region"][worldData["region"] == "Antigua"| worldData['region']=='Barbuda'] <- "Antigua and Barbuda"
-worldData["region"][worldData["region"] == "Palestine"] <- "Palestinian Territory"
-worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Hong Kong'] <- "China, Hong Kong SAR"
-worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Macao'] <- "China, Macao SAR"
-worldData["region"][worldData["region"] == "Guernsey"| worldData['region']=='Jersey'] <- "Channel Islands"
-
-#Merge data
-MergedData <- left_join(worldData, country_total, by = 'region')
+country_total["region"][country_total["region"] == "Cote d'Ivoire"] <- "Côte d'Ivoire"
+country_total["region"][country_total["region"] == "Congo DRC"] <- "Democratic Republic of the Congo"
+country_total["region"][country_total["region"] == "Congo"] <- "Republic of Congo"
+country_total["region"][country_total["region"] == "Cabo Verde"] <- "Cape Verde"
+country_total["region"][country_total["region"] == "Gambia"] <- "The Gambia"
+country_total["region"][country_total["region"] == "China, Hong Kong SAR"] <- "Hong Kong"
+country_total["region"][country_total["region"] == "South Korea"] <- "Republic of Korea"
+country_total["region"][country_total["region"] == "Laos"] <- "Lao PDR"
+country_total["region"][country_total["region"] == "China, Macao SAR"] <- "Macao" 
+country_total["region"][country_total["region"] == "Sao Tome and Principe"] <- "	São Tomé and Principe"
+country_total["region"][country_total["region"] == "North Korea"] <- "Dem. Rep. Korea"
+country_total["region"][country_total["region"] == "Palestinian Territory"] <- "Palestine"
+country_total["region"][country_total["region"] == "Eswatini"] <- "Swaziland"
+country_total["region"][country_total["region"] == "China, Taiwan Province of China"] <- "Taiwan"
+
+# # load map data and align country names
+# worldData <- map_data("world")
+# #correct region names
+# worldData["region"][worldData["region"] == "Russia"] <- "Russian Federation"
+# worldData["region"][worldData["region"] == "USA"] <- "United States"
+# worldData["region"][worldData["region"] == "Virgin Islands" & worldData['subregion']==' US'] <- "US Virgin Islands"
+# worldData["region"][worldData["region"] == "Taiwan"] <- "China, Taiwan Province of China"
+# worldData["region"][worldData["region"] == "Trinidad" | worldData['region']=='Tobago'] <- "Trinidad and Tobago"
+# worldData["region"][worldData["region"] == "Grenadines"| worldData['region']=='Saint Vincent'] <- "Saint Vincent and the Grenadines"
+# worldData["region"][worldData["region"] == "Swaziland"] <- "Eswatini"
+# worldData["region"][worldData["region"] == "UK"] <- "United Kingdom"
+# worldData["region"][worldData["region"] == "Cape Verde"] <- "Cabo Verde"
+# worldData["region"][worldData["region"] == "Republic of Congo"] <- "Congo"
+# worldData["region"][worldData["region"] == "Democratic Republic of the Congo"] <- "Congo DRC"
+# worldData["region"][worldData["region"] == "Ivory Coast"] <- "Cote d'Ivoire"
+# worldData["region"][worldData["region"] == "Brunei"] <- "Brunei Darussalam"
+# worldData["region"][worldData["region"] == "Antigua"| worldData['region']=='Barbuda'] <- "Antigua and Barbuda"
+# worldData["region"][worldData["region"] == "Palestine"] <- "Palestinian Territory"
+# worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Hong Kong'] <- "China, Hong Kong SAR"
+# worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Macao'] <- "China, Macao SAR"
+# worldData["region"][worldData["region"] == "Guernsey"| worldData['region']=='Jersey'] <- "Channel Islands"
+# MergedData <- left_join(worldData, country_total, by = 'region')
+
+world <- ne_countries(scale = "medium", returnclass = "sf")
+world <- world %>% 
+  filter(sov_a3 != 'ATA') #%>% #remove antarctica
+  #select(-c("formal_en","formal_fr","note_adm0","note_brk","name_sort","name_alt"))
+  #select("name_long")
+
+MergedData <- left_join(world, country_total, by = c('name_long'='region'))
 #replace zero with na
@@
 p5 <- ggplot() +
-  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = TotalResults),
-               color = 'black',linewidth=0.15)+
+  #geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = TotalResults),
+  #             color = 'black',linewidth=0.15)+
+  geom_sf(data = MergedData, color = "black",  aes(fill = TotalResults))+
+  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 
+           +datum=WGS84 +units=m +no_defs")+
   scale_fill_fermenter("Total Deaths",
@@
                        label=scales::comma)+
-  theme_void()+
+  theme_bw()+
   theme(legend.spacing.y = unit(.5,'cm'),
@@
         legend.title = element_text(size=13))+
-
-  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
-  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
+  theme(panel.border = element_rect(color = "transparent"),
+        axis.text.x= element_blank(),
+        axis.ticks.x = element_blank())+
+  #theme(panel.grid.major = element_line(color = 'gray75', 
+  #                                      linetype = "solid",
+  #                                      linewidth = 0.25),
+  #      panel.grid.minor = element_line(color = 'gray75', 
+  #                                      linetype = "solid",
+  #                                      linewidth = 0.25),
+  #      panel.grid = element_line(colour = "grey70", size=2),
+  #      )+
+
+  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))#+
+  #coord_proj("+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
+  #coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
 
@@
 
-Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
-input <- file.path(getwd(),'input')
-countryResults <- read.csv(paste0(input,'/BenMAP',"/Country Results Summary by Model & Year.csv"))
+Plots <- file.path( 'analysis','Manuscript','Plots')
+input <- file.path('../..','input')
+
+countryResults <- read.csv(file.path(input,'BenMAP',"Country Results Summary by Model & Year.csv"))
 
@@
 #read in population
-socio_inputs <- file.path('input','RFF','lookup_tables')
+socio_inputs <- file.path('../..','input','RFF','lookup_tables')
 mean.pop<-read.csv(file.path(socio_inputs,"rff_pop_means.csv")) #this data set includes all countries (rff-sp does not)
@@
 
-# read in map data and align country names
-worldData <- map_data("world")
-#correct region names
-worldData["region"][worldData["region"] == "Russia"] <- "Russian Federation"
-worldData["region"][worldData["region"] == "USA"] <- "United States"
-worldData["region"][worldData["region"] == "Virgin Islands" & worldData['subregion']==' US'] <- "US Virgin Islands"
-worldData["region"][worldData["region"] == "Taiwan"] <- "China, Taiwan Province of China"
-worldData["region"][worldData["region"] == "Trinidad" | worldData['region']=='Tobago'] <- "Trinidad and Tobago"
-worldData["region"][worldData["region"] == "Grenadines"| worldData['region']=='Saint Vincent'] <- "Saint Vincent and the Grenadines"
-worldData["region"][worldData["region"] == "Swaziland"] <- "Eswatini"
-worldData["region"][worldData["region"] == "UK"] <- "United Kingdom"
-worldData["region"][worldData["region"] == "Cape Verde"] <- "Cabo Verde"
-worldData["region"][worldData["region"] == "Republic of Congo"] <- "Congo"
-worldData["region"][worldData["region"] == "Democratic Republic of the Congo"] <- "Congo DRC"
-worldData["region"][worldData["region"] == "Ivory Coast"] <- "Cote d'Ivoire"
-worldData["region"][worldData["region"] == "Brunei"] <- "Brunei Darussalam"
-worldData["region"][worldData["region"] == "Antigua"| worldData['region']=='Barbuda'] <- "Antigua and Barbuda"
-worldData["region"][worldData["region"] == "Palestine"] <- "Palestinian Territory"
-worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Hong Kong'] <- "China, Hong Kong SAR"
-worldData["region"][worldData["region"] == "China" & worldData['subregion']=='Macao'] <- "China, Macao SAR"
-worldData["region"][worldData["region"] == "Guernsey"| worldData['region']=='Jersey'] <- "Channel Islands"
-
-#Merge data
-MergedData <- left_join(worldData, country_plot, by = 'region')
+country_plot["region"][country_plot["region"] == "Cote d'Ivoire"] <- "Côte d'Ivoire"
+country_plot["region"][country_plot["region"] == "Congo DRC"] <- "Democratic Republic of the Congo"
+country_plot["region"][country_plot["region"] == "Congo"] <- "Republic of Congo"
+country_plot["region"][country_plot["region"] == "Cabo Verde"] <- "Cape Verde"
+country_plot["region"][country_plot["region"] == "Gambia"] <- "The Gambia"
+country_plot["region"][country_plot["region"] == "China, Hong Kong SAR"] <- "Hong Kong"
+country_plot["region"][country_plot["region"] == "South Korea"] <- "Republic of Korea"
+country_plot["region"][country_plot["region"] == "Laos"] <- "Lao PDR"
+country_plot["region"][country_plot["region"] == "China, Macao SAR"] <- "Macao" 
+country_plot["region"][country_plot["region"] == "Sao Tome and Principe"] <- "	São Tomé and Principe"
+country_plot["region"][country_plot["region"] == "North Korea"] <- "Dem. Rep. Korea"
+country_plot["region"][country_plot["region"] == "Palestinian Territory"] <- "Palestine"
+country_plot["region"][country_plot["region"] == "Eswatini"] <- "Swaziland"
+country_plot["region"][country_plot["region"] == "China, Taiwan Province of China"] <- "Taiwan"
+
+# # load map data and align country names
+world <- ne_countries(scale = "medium", returnclass = "sf")
+world <- world %>% 
+  filter(sov_a3 != 'ATA') #%>% #remove antarctica
+  #select(-c("formal_en","formal_fr","note_adm0","note_brk","name_sort","name_alt"))
+  #select("name_long")
+
+MergedData <- left_join(world, country_plot, by = c('name_long'='region'))
 #replace zero with na
@@
 p5 <- ggplot() +
-  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = death_percap),
-               color = 'black',linewidth=0.15)+
+  geom_sf(data = MergedData, color = "black",  aes(fill = death_percap))+
+  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 
+           +datum=WGS84 +units=m +no_defs")+
   scale_fill_fermenter("Deaths Per Capita",
@@
                         na.value='lightgray',
-                       #values=rescale(c(-1,0,1)),
                        limits=c(#0,
@@
                        labels = c('Minimum',rep('',8),'Maximum'))+
-  theme_void()+
+  theme_bw()+
   theme(legend.position = 'right')+
@@
         legend.title = element_text(size=13))+
-  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
-  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
-
+  theme(panel.border = element_rect(color = "transparent"),
+        axis.text.x= element_blank(),
+        axis.ticks.x = element_blank())+
+  guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))#+
 p5
@@
 p6 <- ggplot() +
-  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, 
-                                      fill = MeanMortalityRate),
-               color = 'black',linewidth=0.15)+
+  geom_sf(data = MergedData, color = "black",  aes(fill = MeanMortalityRate))+
+  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 
+           +datum=WGS84 +units=m +no_defs")+
   scale_fill_fermenter("Respiratory Mortality Rates",
@@
 
-  theme_void()+
+  theme_bw()+
   ggtitle("b) Respiratory Mortality Rates Per 100K")+
@@
   guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
-  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
+  theme(panel.border = element_rect(color = "transparent"),
+        axis.text.x= element_blank(),
+        axis.ticks.x = element_blank())
 p6
@@
 p7 <- ggplot() +
-  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = mean_pop),
-               color = 'black',linewidth=0.15)+
+  geom_sf(data = MergedData, color = "black",  aes(fill = mean_pop))+
+  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 
+           +datum=WGS84 +units=m +no_defs")+
   scale_fill_fermenter("Population",
@@
                        labels = c('Minimum','','','','','','Maximum'))+
-  theme_void()+
+  theme_bw()+
   ggtitle('a) Population')+
@@
   guides(col = guide_legend(override.aes = list(shape = 15, size = 10)))+
-  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
+  theme(panel.border = element_rect(color = "transparent"),
+        axis.text.x= element_blank(),
+        axis.ticks.x = element_blank())
 p7
@@
 p8 <- ggplot() +
-  geom_polygon(data = MergedData, aes(x=long, y = lat, group=group, fill = AvgResponse),
-               color = 'black',linewidth=0.15)+
+  geom_sf(data = MergedData, color = "black",  aes(fill = AvgResponse))+
+  coord_sf(crs = "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 
+           +datum=WGS84 +units=m +no_defs")+
   scale_fill_fermenter("",
@@
                        labels = c('Minimum','','','','','','Maximum'))+
-  theme_void()+
+  theme_bw()+
   ggtitle(expression(paste('c) Average Ozone Response (pptv ', O[3],"/", "ppbv ",CH[4],')')))+
@@
   theme(plot.title = element_text(size = 20))+
-  coord_map(projection = 'mercator',xlim=c(-180,180),ylim = c(-55,80))
+  theme(panel.border = element_rect(color = "transparent"),
+        axis.text.x= element_blank(),
+        axis.ticks.x = element_blank())
 p8
@@
 
-Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
-inputs <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
+Plots <- file.path('../..','analysis','Manuscript','Plots')
+inputs <- file.path('../..','analysis','Manuscript','BenMAP','Outputs')
 countryresults <-
@@
   ungroup() %>%
-  mutate(across(c('means'), ~ signif(.,3)))
+  #mutate(across(c('means'), ~ signif(.,3))) %>%
+  mutate(means = case_when( means >100 ~ signif(means,2), TRUE ~signif(means,1)))
 
@@
   ungroup() %>%
-  mutate(across(c('means'), ~ signif(.,3)))
+  mutate(across(c('means'), ~ signif(.,2)))
 
@@
 #Read in 
-Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
-inputs <- file.path(getwd(),'analysis','Manuscript','BenMAP','Outputs')
-rft_inputs <- file.path(getwd(),'output','npd')
+Plots <- file.path('../..', 'analysis','Manuscript','Plots')
+inputs <- file.path('../..','analysis','Manuscript','BenMAP','Outputs')
+rft_inputs <- file.path('../..','output','npd')
 
@@
   xlab('')+
-  ylab("Change in Global $1660(2020)/mT" ~CH[4])+
+  ylab("Change in Global $1700(2020)/mT" ~CH[4])+
   theme(axis.text = element_text(size=16),

### 2023-05-23 code/3_momm-rft_analysis.R

@@ -1,0 +1,827 @@
+#####
+## Title:   3_momm-rft_analysis.R
+## Purpose: This file will read all output files from the Methane-Ozone Mortality
+##          Model-Reduced Form Tool and generate aggregate dataframes
+##          for use in the manuscript graphics script.
+##          
+## Inputs:  output/damages/rffsp/damages_[scenario].parquet
+## Outputs: output/damages/rffsp/damages_[scenario].parquet
+## Written by: US EPA, Climate Change Division (OAP); April 2023
+## Last updated: 4/12/2023 by E. McDuffie
+## Notes:       
+##
+#####
+
+###### Clear Workspace ######
+rm(list = ls()); gc()
+
+###### Load packages ###### 
+require(tidyverse);
+require(pbapply);
+require(arrow);
+require(magrittr);
+require(dplyr);
+
+
+###### Set Paths ######
+#eem (b/c I don't have permission to open SCC_FrEDI.Rproj)
+#setwd("/home/emcduf01/shared/ncee/scghg/FrEDI/SCC_FrEDI/")
+inputsPath   <- "." %>% file.path("output","rft")
+outputsPath  <- "." %>% file.path("output", "summary_data")
+#original scghg
+#scghgPath    <- "." %>% file.path("output","scghg")
+##transformed scghg path (updated 1/5/23)
+#scghgPath    <- "." %>% file.path("output","scghg_damage_transformation")
+
+##### Collect Damage Files ####
+c_iteration  <- inputsPath %>% list.files(pattern = "\\parquet") %>%
+  (function(x){sub("\\_momm_rft.parquet", "", x)}) %>%
+  (function(x){sub("damages_", "", x)}) %>%
+  as.numeric %>% sort; 
+c_iteration %>% length
+#c_iteration <- c_iteration[1:10]
+
+#excluded_sectors <- c('Extreme Temperature','CIL Extreme Temperature','Asphalt Roads')
+#c_select_rawCols <- c("sector", "variant", "impactType", "region", "year", "model_type", 
+#                      "sectorprimary", "physicalmeasure", "trial", "damageType","driverType","driverValue","reg_pop","national_pop","gdp_usd")
+
+#### Set Progress Bar ####
+op <- pboptions(type = "timer",char = "=")
+
+
+
+###### Read-in/Load Data ######
+
+###### Aggregate All Trial Results ######
+##1. Read in all physical and annual damages from all trials
+# If df is too large...
[…]

### 2023-05-23 code/BenMAP/BenMap_O3Mort_Valuation.R

@@ -1,0 +1,457 @@
+#####
+## Title:   BenMAP_O3Mort_Valuation.R
+## Purpose: This file reads in O3 Mortality results from BenMAP
+##          and calculates the annual damages$ and 
+##          calculates the NPD value ($ per ton CO2) using 
+##          1.5%, 2%, 2.5%, and 3% Ramsey discounting and a 2% and 3% constant discount rate
+##          by 2100 in 2020 USD.
+## Inputs:  
+## Outputs: output/npd/co2_full_streams_2020_national_default_adaptation.parquet
+##          output/npd/npd_fredi_national.csv
+## Written by: US EPA, Climate Change Division (OAP); March 2023
+## Last updated: 3.30.23 by E. McDuffie
+## NOTES: 
+#####
+
+##########################
+#################  library
+##########################
+
+## clear workspace
+rm(list = ls())
+gc()
+
+## this function will check if a package is installed, and if not, install it
+list.of.packages = c('tidyverse', 'magrittr',
+                     'arrow',
+                     'zoo',
+                     'doParallel', 'foreach',
+                     'ggplot2')
+new.packages = list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages, repos = "http://cran.rstudio.com/")
+lapply(list.of.packages, library, character.only = TRUE)
+
+##########################
+################### constants
+##########################
+
+
+Inputs_mort <- file.path('Results','Global BenMAP WebTool')
+Inputs_socio <- file.path('Code','InputData','Tool Inputs','RFF') #need mean population that was used in BenMAP, as well as the corresponding GDP scenario
+cntryxwalkpath <- file.path('Code/InputData/Tool Inputs')
+Results <- file.path('Results')
+
+#### Constants
+base_vsl = 9.52e6 #us vsl in 2020$ 
+## WHAT VSL TO USE?? ##
+# EPA 2010 (updated 2014) guidance = $7.9 million in 2008$ (= 9.52mil in 2020$)
+# What is used in FrEDI?
[…]

### 2023-08-11 analysis/Manuscript/BenMAP/BenMap_O3Mort_Valuation.R
Updates to code organization

@@ -1,369 +1,0 @@
-##################################################################
-## Title:   BenMAP_O3Mort_Valuation.R
-## Purpose: This file reads in O3 Mortality results from BenMAP
-##          and calculates the annual damages$ and 
-##          calculates the NPD value ($ per ton CO2) using 
-##          1.5%, 2%, 2.5%, and 3% Ramsey discounting and a 2% and 3% constant discount rate
-##          by 2100 in 2020 USD.
-##          Option to implement cessation lags or not
-## Inputs:  Final Country Grid Col_Row Index.csv
-##          rffsp_pop_gdp_all_trials.parquet
-##          Country Results Interpolated 2020-2100 by Model.csv
-## Outputs: Country Results Interpolated Mortality Annual Undiscounted Damages_w[out]Cessation.csv
-##          npd_damage_transformation_full_streams_w[out]Cessation.parquet
-##          npd_global_damage_transformation_w[out]Cessation.csv
-## Written by: US EPA, Climate Change Division (OAP); March 2023
-## Last updated: 5.11.23 by E. McDuffie
-## NOTES: 
-#####################################################################
-
-##########################
-#################  library
-##########################
-
-## clear workspace
-rm(list = ls())
-gc()
-
-## this function will check if a package is installed, and if not, install it
-list.of.packages = c('tidyverse', 'magrittr',
-                     'arrow',
-                     'zoo',
-                     'doParallel', 'foreach',
-                     'ggplot2')
-new.packages = list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-if(length(new.packages)) install.packages(new.packages, repos = "http://cran.rstudio.com/")
-lapply(list.of.packages, library, character.only = TRUE)
-
-##########################
-################### constants
-##########################
-
-
-Inputs_mort <- file.path('analysis','Manuscript','BenMAP','Outputs')
[…]

### 2023-08-11 code/2_Global Ozone Reduce Form Tool.R

@@ -114,14 +114,15 @@
   NOx_scalar = 1          #Options: numerical values > 0 (e.g., 50% = 0.5, 150% = 1.5)
   # this value represents the proportion of NOx emissions in each country relative
   # to NOx emission in the original CCAC simulations
+  # default = 1
   
-  # Implement cessation lags (1 = yes, 0 = no)
-  cessation_flag = 1
+  # Implement cessation lags (1 = yes, 0 = no (default))
+  cessation_flag = 0
   
   #homogeneous methane-ozone response
-  homogeneous =1
+  homogeneous = 0
   # if set to 1, the ozone response will be set to the global (non-weighted) average (4.8 pptv/ppbv)
-  
+  #default= 0
   
 ##########
 ###### Pre-Processing Steps

### 2023-08-11 code/BenMAP/BenMap_O3Mort_Valuation.R

@@ -1,17 +1,21 @@
-#####
+##################################################################
 ## Title:   BenMAP_O3Mort_Valuation.R
 ## Purpose: This file reads in O3 Mortality results from BenMAP
 ##          and calculates the annual damages$ and 
 ##          calculates the NPD value ($ per ton CO2) using 
 ##          1.5%, 2%, 2.5%, and 3% Ramsey discounting and a 2% and 3% constant discount rate
 ##          by 2100 in 2020 USD.
-## Inputs:  
-## Outputs: output/npd/co2_full_streams_2020_national_default_adaptation.parquet
-##          output/npd/npd_fredi_national.csv
+##          Option to implement cessation lags or not
+## Inputs:  Final Country Grid Col_Row Index.csv
+##          rffsp_pop_gdp_all_trials.parquet
+##          Country Results Interpolated 2020-2100 by Model.csv
+## Outputs: Country Results Interpolated Mortality Annual Undiscounted Damages_w[out]Cessation.csv
+##          npd_damage_transformation_full_streams_w[out]Cessation.parquet
+##          npd_global_damage_transformation_w[out]Cessation.csv
 ## Written by: US EPA, Climate Change Division (OAP); March 2023
-## Last updated: 3.30.23 by E. McDuffie
+## Last updated: 5.11.23 by E. McDuffie
 ## NOTES: 
-#####
+#####################################################################
 
 ##########################
 #################  library
@@ -36,110 +40,54 @@ lapply(list.of.packages, library, character.only = TRUE)
 ##########################
 
 
-Inputs_mort <- file.path('Results','Global BenMAP WebTool')
-Inputs_socio <- file.path('Code','InputData','Tool Inputs','RFF') #need mean population that was used in BenMAP, as well as the corresponding GDP scenario
-cntryxwalkpath <- file.path('Code/InputData/Tool Inputs')
-Results <- file.path('Results')
+Inputs_mort     <- file.path('analysis','Manuscript','BenMAP','Outputs')
[…]

### 2023-08-13 analysis/Manuscript/Paper_RFigures.Rmd
paper review updates

@@
 library('reshape2')
-library('patchwork')
-library('RColorBrewer')
+require('patchwork')
+require('RColorBrewer')
 #library('maps')
@@
 library('arrow')
-library('readxl')
+require('readxl')
 library('zoo')
@@
 #define paths
-Plots <-file.path('analysis','Manuscript','Plots') 
-inputs <- file.path("analysis","Manuscript")
-inputs2 <- file.path('input','RFF','lookup_tables')
-inputs3 <- file.path("input")
-Inputs_socio <- file.path('input','RFF','rft_inputs')
+Plots <-file.path('../..','analysis','Manuscript','Plots') 
+inputs <- file.path('../..',"analysis","Manuscript")
+inputs2 <- file.path('../..','input','RFF','lookup_tables')
+inputs3 <- file.path('../..',"input")
+Inputs_socio <- file.path('../..','input','RFF','rft_inputs')
 
@@
 # Pre-processed in /code/1_build_inputs_for_momm_rft.R
-rffsp <- read_parquet(file.path(getwd(),Inputs_socio, 'rffsp_pop_gdp_all_trials.parquet'))
+rffsp <- read_parquet(file.path(Inputs_socio, 'rffsp_pop_gdp_all_trials.parquet'))
 
@@
 ggsave(plot = p1_bounds, filename = 
-         file.path(Plots,'Plots/Timeseries_O3Mort_CRFbounds.png'),width = 9,height = 5)
+         file.path(Plots,'Fig2a.png'),width = 9,height = 5)
 
@@
 #map of country mortality results
-MODEL = 'GFDL'
-
-Plots <- file.path(getwd(), 'analysis','Manuscript','Plots')
-input <- file.path(getwd(),'input','BenMAP')
+MODEL = 'MMM'
+
+Plots <- file.path('../..','analysis','Manuscript','Plots')
+input <- file.path('../..','input','BenMAP')
 countryResults <- read.csv(paste0(input,"/Country Results Summary by Model & Year.csv"))
@@
 
-Plots <- file.path( 'analysis','Manuscript','Plots')
+Plots <- file.path('../..','analysis','Manuscript','Plots')
 input <- file.path('../..','input')
@@
 countryresults <-
-   read_csv(file.path(inputs,"npd_country_damage_transformation_woutCessation.csv"),
+   read_csv(file.path(inputs,"npd_country_damage_transformation_woutCessation_vsl10.csv"),
            col_types = cols())
@@
 globalresults <-
-  read_csv(file.path(inputs, "npd_global_damage_transformation_woutCessation.csv"),
+  read_csv(file.path(inputs, "npd_global_damage_transformation_woutCessation_vsl10.csv"),
            col_types = cols())
@@
 Results <- file.path('Results')
+# CRF results - from BenMAP valuation script
 globalresults <-
-  read_csv(file.path(inputs,"npd_global_damage_transformation_woutCessation.csv"),
+  read_csv(file.path(inputs,"npd_global_damage_transformation_woutCessation_vsl10.csv"),
            col_types = cols())
+
+#Socioeconomic results - from RFT code
 Results2 <- file.path(rft_inputs,'npd_global_rff_means_NOx_1_MMM.csv')
 rft_results <- read_csv(Results2, col_types = cols())
-nox_results1 <- file.path(rft_inputs,'npd_global_means_NOx_1_MMM.csv')
+
+#NOx sensitivity results- from RFT code
+nox_results1 <- file.path(rft_inputs,'npd_global_means_vsl10_NOx_1_MMM.csv')
 nox_results1 <- read_csv(nox_results1, col_types = cols())%>%
   mutate(NOx = '1')
-nox_results5 <- file.path(rft_inputs, 'npd_global_means_NOx_0.5_MMM.csv')
+nox_results5 <- file.path(rft_inputs, 'npd_global_means_vsl10_NOx_0.5_MMM.csv')
 nox_results5 <- read_csv(nox_results5, col_types = cols())%>%
   mutate(NOx = '0.5')
-nox_results15 <- file.path(rft_inputs, 'npd_global_means_NOx_1.5_MMM.csv')
+nox_results15 <- file.path(rft_inputs, 'npd_global_means_vsl10_NOx_1.5_MMM.csv')
 nox_results15 <- read_csv(nox_results15, col_types = cols()) %>%

### 2023-08-13 code/2_Global Ozone Reduce Form Tool.R

@@ -3,7 +3,7 @@
 ### Created by: Melanie Jackson, IEc 
 ### Adapted by: E. McDuffie (EPA)
 ### Date Created: 3/15/2023      
-### Last Edited: 4/24/2023
+### Last Edited: 8/11/2023
 ##      Inputs:                                                                       ##
 ##        -Country Results Interpolated 2020-2100 by Model.csv                        ##
 ##        -All Trajectory Mortality Ratios_through 2100.rds (parts 1&2)               ##
@@ -20,7 +20,7 @@
   rm(list = ls()); gc()
   
   #necessary packages
-    packs<-c("dplyr","tidyverse","readxl","purrr","foreach","utils","pbmcapply")
+    packs<-c("dplyr","tidyverse","readxl","purrr","foreach","utils","pbmcapply","arrow")
     for (package in packs) { #Installs packages if not yet installed
       if (!requireNamespace(package, quietly = TRUE))
         install.packages(package)
@@ -51,7 +51,7 @@
   BaseYear            = 2020 # year of CH4 pulse used in BenMAP simulations
   gdp_2011_to_2020    = 113.784/98.164 # GDP Implicit Price Deflators (https://apps.bea.gov/iTable/?reqid=19&step=3&isuri=1&select_all_years=0&nipa_table_list=13&series=a&first_year=2006&last_year=2020&scale=-99&categories=survey&thetable= )
                       #last access: March 30, 2023
-  base_vsl            = 9.33e6  # = USD VSL in 2006$, inflated to 2020$ (from EPA 2010)
+  base_vsl            = 10.05e6 # $2020 VSL from Rennert et al., 2022 (and EPA SC-GHG report) 9.33e6  # = USD VSL in 2006$, inflated to 2020$ (from EPA 2010)
   Elasticity          = 1  # income elasticity
   Years               <- c(2020:2100) # set simulation years
   lags                = c(6, 2.5, 2.5, 2.5, 2.5, 4/15, 4/15, 4/15, 4/15,4/15, 4/15, 4/15, 4/15, 
@@ -65,32 +65,13 @@
 ########
     
     ## 1) Original BenMAP O3-Methane Mortality 
-    CloudResults <- read.csv(file.path(Inputs,"BenMAP","Country Results Interpolated 2020-2100 by Model_eem.csv"))
+    CloudResults <- read.csv(file.path(Inputs,"BenMAP","Country Results Interpolated 2020-2100 by Model.csv"))
[…]

### 2023-08-13 code/3_methane_ozone_npd.R

@@ -5,9 +5,9 @@
 ##                and aggregate statistics across all trials. (for each model & country)
 ##                The calculations are done for each model and country and combined at
 ##                the end to help reduce the memory usage when processing all 10,000 RFF scenarios
-### Written by:   US EPA, Climate Change Division (OAP), adapted from NCEE (OP)
+### Written by:   US EPA, Climate Change Division (OAP)
 ### Date Created: 3/15/2023      
-### Last Edited:  4/25/2023
+### Last Edited:  8/13/2023
 ##      Inputs:                                                                       ##
 ##        -damages_mean_NOx_X_momm_rft.parquet                                        ##
 ##        -Final Country Grid Col_Row Index.csv                                       ##
@@ -41,7 +41,7 @@ Outputs <- file.path("output",'npd')
 
 YEAR  = 2020
 PULSE = 275e6 #CH4 pulse of 275 million metric tonnes
-allRFF = 1
+allRFF = 1 #(set to 1 if reading in all RFF scenario, otherwise read in mean)
 
 countries <- read.csv(file.path('input',"Final Country Grid Col_Row Index.csv"))[,c(1,3:4,6,8)]
 countries <- countries %>% filter(Region != "")
@@ -72,12 +72,12 @@ for (MODEL in c('MMM')) { #},'CESM2','HadGEM','GISS','GFDL','MIROC')) {
   if (allRFF==1){
     # To read all files:
     damages = 
-      list.files(Inputs, full.names = T) %>% 
+      list.files(Inputs, pattern = "damages_\\d+\\_vsl10_NOx_1*", full.names = T) %>% 
       map_df(~read_momm_rft(., MODEL))
   } else {
     # to read specific file:
     damages =
-      read_momm_rft(file.path(Inputs,paste0('damages_mean_NOx_',NOx_scalar,'_momm_rft.parquet')),MODEL)
+      read_momm_rft(file.path(Inputs,paste0('damages_mean_vsl10_NOx_',NOx_scalar,'_momm_rft.parquet')),MODEL)
   }
   print(MODEL)
   
@@ -99,10 +99,9 @@ for (MODEL in c('MMM')) { #},'CESM2','HadGEM','GISS','GFDL','MIROC')) {
[…]

### 2023-08-13 code/BenMAP/BenMap_O3Mort_Valuation.R

@@ -13,7 +13,7 @@
 ##          npd_damage_transformation_full_streams_w[out]Cessation.parquet
 ##          npd_global_damage_transformation_w[out]Cessation.csv
 ## Written by: US EPA, Climate Change Division (OAP); March 2023
-## Last updated: 5.11.23 by E. McDuffie
+## Last updated: 8.11.23 by E. McDuffie
 ## NOTES: 
 #####################################################################
 
@@ -44,13 +44,13 @@ Inputs_mort     <- file.path('analysis','Manuscript','BenMAP','Outputs')
 Inputs_socio    <- file.path('input','RFF') #need mean population that was used in BenMAP, as well as the corresponding GDP scenario
 cntryxwalkpath  <- file.path('input')
 Results         <- file.path('analysis','Manuscript','BenMAP','Outputs')
-
+  
 #### Constants
-file_name_flag  <- 'vsl_10'  #custom file name string ending
+file_name_flag  <- 'vsl10_elas14'  #custom file name string ending
 pulse           = 275e6     #original methane pulse was 275 million metric tonnes
-base_vsl        = 10.05e6#9.33e6    #us vsl in 2020$ 
+base_vsl        = 10.05e6 # US $2020 VSL (from Rennert et al., 2022/EPA SC-GHG Report) 9.33e6    #us vsl in 2020$ 
 # from USEPA 2006 value in 2010 guidance (default = 1)
-Elasticity      = 1
+Elasticity      = 1.44
 # turn 20-year cessation flag on (=1) or off (=0) (default = 0)
 cessation_flag  = 0
 

### 2023-08-14 analysis/Manuscript/Paper_RFigures.Rmd
Figure 3 updates

@@
 
-  # Figure SX
+  # Figure S7
 
@@
         scale_x_discrete(labels = function(x) str_wrap(x,width=30))+
-        scale_y_continuous(limits=c(0,3000), label = scales::dollar)+
+        scale_y_continuous(limits=c(0,3200), label = scales::dollar)+
         xlab("")+ylab("$(2020)/mT" ~CH[4])+
@@
               axis.title = element_text(size=16))
+  p1
+
 # Figure 2c
@@
 #Socioeconomic results - from RFT code
-Results2 <- file.path(rft_inputs,'npd_global_rff_means_NOx_1_MMM.csv')
+Results2 <- file.path(rft_inputs,'npd_global_rff_means_vsl10_NOx_1_MMM.csv')
 rft_results <- read_csv(Results2, col_types = cols())
@@
   xlab('')+
-  ylab("Change in Global $1700(2020)/mT" ~CH[4])+
+  ylab("Change in Global $1800(2020)/mT" ~CH[4])+
   theme(axis.text = element_text(size=16),
         axis.title = element_text(size=16))+
-  scale_y_continuous(limits = c(-1000, 1300),label = scales::dollar)+
+  scale_y_continuous(limits = c(-1100, 1300),label = scales::dollar)+
   theme(legend.position = c(0.8,0.25),

### 2023-08-14 code/3_momm-rft_analysis.R

@@ -1,827 +1,0 @@
-#####
-## Title:   3_momm-rft_analysis.R
-## Purpose: This file will read all output files from the Methane-Ozone Mortality
-##          Model-Reduced Form Tool and generate aggregate dataframes
-##          for use in the manuscript graphics script.
-##          
-## Inputs:  output/damages/rffsp/damages_[scenario].parquet
-## Outputs: output/damages/rffsp/damages_[scenario].parquet
-## Written by: US EPA, Climate Change Division (OAP); April 2023
-## Last updated: 4/12/2023 by E. McDuffie
-## Notes:       
-##
-#####
-
-###### Clear Workspace ######
-rm(list = ls()); gc()
-
-###### Load packages ###### 
-require(tidyverse);
-require(pbapply);
-require(arrow);
-require(magrittr);
-require(dplyr);
-
-
-###### Set Paths ######
-#eem (b/c I don't have permission to open SCC_FrEDI.Rproj)
-#setwd("/home/emcduf01/shared/ncee/scghg/FrEDI/SCC_FrEDI/")
-inputsPath   <- "." %>% file.path("output","rft")
-outputsPath  <- "." %>% file.path("output", "summary_data")
-#original scghg
-#scghgPath    <- "." %>% file.path("output","scghg")
-##transformed scghg path (updated 1/5/23)
-#scghgPath    <- "." %>% file.path("output","scghg_damage_transformation")
-
-##### Collect Damage Files ####
-c_iteration  <- inputsPath %>% list.files(pattern = "\\parquet") %>%
-  (function(x){sub("\\_momm_rft.parquet", "", x)}) %>%
-  (function(x){sub("damages_", "", x)}) %>%
-  as.numeric %>% sort; 
-c_iteration %>% length
-#c_iteration <- c_iteration[1:10]
-
-#excluded_sectors <- c('Extreme Temperature','CIL Extreme Temperature','Asphalt Roads')
-#c_select_rawCols <- c("sector", "variant", "impactType", "region", "year", "model_type", 
-#                      "sectorprimary", "physicalmeasure", "trial", "damageType","driverType","driverValue","reg_pop","national_pop","gdp_usd")
-
-#### Set Progress Bar ####
-op <- pboptions(type = "timer",char = "=")
-
-
-
-###### Read-in/Load Data ######
-
-###### Aggregate All Trial Results ######
-##1. Read in all physical and annual damages from all trials
-# If df is too large...
[…]
